### 2016-04-25 linvis.tex
Adding Related works

% vim: set fenc=utf-8 ft=latex encoding=utf-8
% -*- mode: latex; coding: UTF-8; -*-

\newif\ifdraft
%\drafttrue

\ifdraft
        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
        \def\baselinestretch{1}
        \setlength{\marginparwidth}{2cm}
\else
        \documentclass[conference]{IEEEtran}
\fi

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\newcommand{\TheTitle}{Visualizing Release Information of Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

\synctex=1

\usepackage[hyphens]{url}
\urlstyle{same}

\ifdraft
        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
                pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}

\else
        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
                pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
\fi

\usepackage[nospace]{cite}

% Table Support
\usepackage{dcolumn}
\usepackage{longtable}

\usepackage{balance}
\usepackage{placeins}
\usepackage{multirow}

% Extra support
\usepackage{xspace}
\usepackage{caption}
\usepackage[nospace]{cite}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Fix any bad-hyphenations here
\hyphenation{contained}

\input{macros.tex}

\begin{document}

\title{\TheTitle}
\author{
\IEEEauthorblockA{\TheAuthors}
\IEEEauthorblockN{Department of Computer Science,
                    University of Victoria, Canada.}
\IEEEauthorblockA{Email: \TheEmails}
}
\maketitle
\begin{abstract}

        With an average of more than 900 top-level merges into the Linux kernel per
        release, many containing hundreds of commits and some containing thousands,
        maintenance of older versions of the kernel becomes nearly impossible.
        Various commercial products, such as the Android platform, run older
        versions of the kernel. Due to security, performance, and changing hardware
        needs, maintainers must understand what changes (commits) are added the current version of the kernel 
        since the last time they inspected it in order to make the necessary
        patches.

        Current tools provide information about repositories through the directed
        acyclic graph (DAG) of the repostory, which is helpful for smaller projects. However, with
        the scale and number of branches in the kernel the DAG becomes overwhelming
        very quickly. Furthermore, the DAG contains every ancestor of every commit, while maintainers are 
        more interested in how and when a commmit arrives to Linux official repository.

        In this paper, we present a tool that uses a tree-based model, as described
        by German et al., to construct a clearer summary of the changes made. This
        tool is designed to allow users to explore the topology of the repository
        for a given release of the kernel and view an aggregated summary of the
        changes made.

\end{abstract}

\begin{IEEEkeywords}
\TheKeywords
\end{IEEEkeywords}

\section{Introduction}

Maintainers of older versions of the Linux kernel (such as those used in embedded systems and mobile phones) must keep track of changes to the current version
of the current in order to determine what commits are relevant to their versions.
Due to the ratio of growth of the kernel (every year, between 50 and 70 thousand commits get added to it)
this process requires
sifting through thousands of
commits with tools that are unable to filter and effectively visualize projects at
the scale of the kernel. Tools like Gitk use a directed acyclic graph (DAG) model,
showing all commits and merges in chronological order by when the commit was
authored, and not by then it arrives in the official repository.

This representation works in smaller projects; it enables users to see
when changes are made, when these changes are merged, and how each branch is
interacting.  
In large modular projects, like the Linux kernel, the DAG becomes a
mess of merges and commits (figure \ref{fig:gitk}) losing its visual meaning.  In
some cases, the Linux kernel is simply too large for the system to generate a
visualization (eg. Github provides a DAG view for many projects, but is unable to
display the visualization for projects as large as the Linux kernel (figure
\ref{fig:gitfail}).
Furthermore, Linux creates between 60 and 70k new commits every year and, according to our previous work \cite{German2015}pdfl, 
a commit takes a median of 30 days (from the moment it is authored) to the time it arrives to the official repository
This means that, if a person inspected the commits in this repository today, tomorrow it might contain  many more
of commits that were authored long before. However, distinguishing these new commits is not trivial.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitk.png}
        \caption{A sample DAG view of commits from Linux in Gitk}
        \label{fig:gitk}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project}
        \label{fig:gitfail}
\end{figure}

One of the challenges with visualizing the arrival of commits to a repository is that
in Git, commits are immutable, and Git does not store the date that a commit was merged into another branch (including master).
To complicate the problem, the DAG points to the ancestors of a commit (a model necessary for the operation of Git), but developers 
would prefer to know the path that a commit followed to reach the master repository. \dmg{perhaps a figure is good here}.
Hence, tracing the path that any commit followed to master would imply that for any given merge, it would be possible to know what commits were merged. A user could
inspect the commits that have arrived into master within a given time frame by simply checking which commits where merged during that time frame.

% \dmg{I am going to describe the method we used here... }

% German et al. recorded the metadata of every commit and merge in the Linux git
% repository starting in 2011 \cite{German2015}, including chronological data that is
% not stored by git. We are able to use this information to determine when a commit
% has been merged into the kernel, regardless of rebases preceding and proceeding that
% point in time. This information is then converted by a heuristic method, described
% by German, to generate sets of trees, which we use as our core model instead of
% using the DAG. Each tree is rooted by the top-level merge performed by Linus, which
% merges the branches or a set of commits into the kernel.

 The contributions of this paper are two. First, we describe a method to convert the DAG of the Linux repository into a tree (we call this the \emph{tree-model of the
 repository}) that represents the path used by
 any commit to reach master. Second, 
 we present a method to inspect and visualize the history of merges in the  Linux project using the tree-model of the repository

These methods and visualizations have been implemented in a web-based
tool\footnote{Our visualization tool is currently available at \url{http://li.turingmachine.org}}
instead of the DAG. Our visualizations and tool provide information about the location of any given
commit in the respective merge tree, the files edited, the modules edited, and the
commit message for a given commit or merge. Our tool allows users to apply various
filters, including the release version, along with a keyword or phrase from the log
preview, the name of the author, or the commit ID. The user can request all the
root-level merges containing a commit or merge that matches the query, or all
commits and merges that match the query.

\evantodo{move to discussion}

Using the tree provides many advantages to the DAG model, but there are two
drawbacks. The first drawback is that a server must query the state of the master
repository since time information is not stored by git. The second drawback is that
the tree construction is a heuristic approach, and in some cases may break.

Our core model provides advantages to the DAG model used by other tools. We separate
the commits and merges into sub-trees and focus on a single sub-tree at a time. This
enables us to filter out commits and merges that do not pertain to the component of
the kernel that we are interested in, providing a clearer picture of what is
happening in a given merge. The tree model also enables us to aggregate metadata
easily. With the addition of the continuous polling of the repository, the model
provides us with a mechanism of performing date-range queries with meaningful
results.

\section{Merge-Tree model}

Git uses a Directed Acyclic Graph (DAG) as its main data model. In this model, a commit has one or more ordered parents (ancestors), except the root commits of a repository which
do not have ancestors--Linux git repository has two such commits. Commits are divided into merge commits (they merge two or more branches) and non-merge
commits. Non-merge commits have only one parent, while merge commits have two or more. The order of the parents matter: the first parent is the branch in which
the merge is being done; while the rest indicate the branches being merged (git allows any number of branches to be merged at the same time). Their order in the
parents relationship indicate the order in which the branches are being merged (second parent first, third parent second, etc.).

In Git, commits are immutable. Once they are created, they never change. Git allows operations to alter commits or to reorder them. However, in such case, the
commit changes its commit id, effectively becoming a new commit. This means that commits do not record the branches or repositories they have traversed in their
way to master. 
For instance, assume the sequence of commits  exemplified in Figure~\ref{fig:repoEvents}. In this case commits are performed in various repos and branches. The
DAG representation of these commits is shown in Figure~\ref{fig:repoDAG}. Notice that this DAG loses information about what repository is Master (in fact, it
loses all information about what repositories any commit has passed through). The merge-tree view of this DAG is shown in Figure~
\ref{fig:repoTree}. Note that the direction of the edges of the DAG has been inverted. Instead of pointing to the ancestors, it points to its successors (the path to
Master). Also, the DAG has been simplified, and it only shows for any commit, one edge (its path to Master).

\begin{figure}[htbp]
  \centering
        \includegraphics[width=\columnwidth]{figures/events.pdf}
  \caption{Example of a sequence of events performed in different repositories. The horizontal axis represents time. Each horizontal section represents a
    different branch and/or repository. Each commit points to its ancestor.}
  \label{fig:repoEvents}
\end{figure}

\begin{figure}[htbp]
  \centering
        \includegraphics[width=.8\columnwidth]{figures/dag.pdf}
  \caption{DAG representation of the commits represented in Figure~\ref{fig:repoEvents}. It loses information about what repository the commit is performed and
    through which repositories it has passed in its way to Master. In fact, the DAG does not even know what Master is.}
  \label{fig:repoDAG}
\end{figure}

\begin{figure}[htbp]
  \centering
        \includegraphics[width=\columnwidth]{figures/tree.pdf}
  \caption{Merge-tree view of the commits  represented in Figure~\ref{fig:repoEvents} showing the path they followed to reach master. In this model the successors of
    each commit represents the path followed by that commit to reach Master.}
  \label{fig:repoTree}
\end{figure}

\subsection{Computing the merge-tree of the DAG of Linux}

Computing the tree-model of a DAG of any repository might not be possible. However, certain features of the development process of Linux make it feasible to
compute the its merge-tree model. In particular, the Master branch of Linux is maintained by Linus Torvalds and only him has write access to it. We have verified this
assertion if previous research~\cite{German2015}. We have developed a heuristic that is presented in Algorithm~\ref{fig:alg}. In a nutshell, first identify the commits directly made
to the master branch. Then,
recursively, for
every commit, using the DAG, determine the shortest path (in terms of time) from such commit to master using the inverted DAG.

\begin{algorithm}
\caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
\begin{algorithmic}
\Function{ComputeTreeModel}{DAG}: tree 
\State {\# Compute the tree from the DAG of a git repository.}
\State {\#  Returns $Tree$, a graph containing every commit }
\State {\# in DAG with the path it followed to master.\\}

\State $head \gets \textit{Head of master of git repository}$
\State $master \gets \textit{traverse DAG from head using }$
\State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
\State $nodes(Tree) \gets nodes(DAG)$
\State \Function{distance2Master}{cid} : seconds 
\State {\# Compute the shortest distance to master setting}
\State {\# cid's successor (next) in its way to master.}
\State {\# This function should be memoized. Otherwise it}
\State {\# would run in exponential time.}
\If {\textit{cid in master}} 
\State \Return 0
\EndIf
\State    $d \gets 	\infty$
   \State {\# Traverse the inverted DAG}
   \For{$c \in children(cid, DAG)$} 
      \If {$c \in master$} 
      \State $d_1 \gets commitTime(c)-commitTime(cid)$
      \Else
      \State {$d_1 \gets distance2Master(c)$} 
      \EndIf
    \If {$d_1 < d $} 
    \State $next \gets c$
   \State  $d \gets d_1$
   \EndIf
   \EndFor
   \State {\# $c$ is the commit that follows $cid$}
   \State {\# in its way to master}
   \State add edge $(cid, next)$ to $Tree$
   \State \Return $d$ 
\EndFunction

\State {\# Compute the distance for each commit}
\State {\# discarding result}
\For{$c \in nodes(DAG)$}
\State $distance2Master(c)$
\EndFor
\State \Return $Tree$
\EndFunction
\end{algorithmic}
\end{algorithm}

\subsection{Evaluation}

Merges that do not have conflicts provide information to verify this heuristic. If a merge does not contain a conflict, it records a summary of the commits that
it merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a list of the first 20 non-merge commits in the merge, including their one-line
log description, the full logs of the merge commits that merge this subset, and the total number of non-merge commits in the merge.

\begin{figure}[htbp]
  \centering
{\tiny
\begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

    Merge branch 'for-linus' of git://neil.brown.name/md

    * 'for-linus' of git://neil.brown.name/md: (24 commits)
      md: clean up do_md_stop
[... edited for the sake of space]
      md: split out md_rdev_init
      md: be more careful setting MD_CHANGE_CLEAN
      md/raid5: ensure we create a unique name for kmem_cache when mddev has no gendisk
      ...
\end{verbatim}}
  \caption{Example of how merges record a subset of commits being merged. The commit only shows the first 20 one line log messages for the 24 non-merge commits
    it merged. The ending ... is part of the log and represents that other commits were merged.}
  \label{fig:sampleMerge}
\end{figure}

We use this information to evaluate the accuracy of the merge tree-model extracted from the DAG. We therefore compare, for commits with no conflict: a) the number of
non-merge commits, b) the summaries of the commits available in the log, and c) the summaries of the merges included.

The results of this verification show that this heuristic detected no incorrect merges since 2013, two incorrect merges in 2012, and two in 2011. However, due
to fast-forward commits performed in Linus Torvalds repository, this heuristic breaks apart on Oct 15, 2009, and no longer works for older merges.

\section{Related Work}

\evantodo{pick 'gitk' or 'Gitk'}

Version control systems monitor the development lifetime of software projects. This
makes the version control system vital in providing information about how a software
project is being developed. To our knowledge, we know of no git repository
visualization tool that builds a tree that maps the path a commit follows to the
master branch from the DAG provided by git. This may be because more information is
required to generate the tree-model than what is stored by git. However, there has
been a lot of work in providing visualizations of various repositories.

Many of the tools work to address the issues of communication between developers and
inter-team collaboration. Codebook uses a mining technique to determine the
developer of a piece of code, the program manager who wrote the specification for
the code, and the program managers and developers on the team who were working
together\cite{Begel2010}. This is designed to help collaboration between developers
on a project while the project is being developed.  In maintaining the code, the
code has already been written, the original developer many not have seen this piece
of code in years, and will most likely have moved to another problem. If the
maintainer is pulling from a part of the kernel that has been recently updated, in
the event of a zero-day attack for instance, the information about who is developing
that portion of the kernel is of use; however, if the maintainer is pulling from
sections that are older, then the maintainer must be able to understand the code
using the provided tools, not the original developer.

Hipikat is another tool with a focus on communication issues between developers. It
is designed for non-collocated teams, helping new developers understand design
decisions of the system \cite{Cubranic2005}. It records artifacts about the project
to build a "memory" of the project development that new users can work with to
understand where and how to make modifications in large projects. The artifacts are
recorded using tools like bugzilla and CVS. Such a tool could prove useful in
understanding the changes between a maintainer's current version of an older kernel
and the changes being made to the current master version of the kernel.

These tools work toward assisting the communications between developers and
providing a summary of the project. Our goal includes providing a visual summary of
the topology of the repository, as well as providing a summary of the changes made
at a location within the repository. EPOSee, Gource, and Gitk are all tools designed
for visualizing various aspects of repositories.

EPOSee is a tool for visualizing changes that often happen simultaneously in a
commit or as a sequence of events \cite{Burch2005}. It produces visualizations about
the confidence that changes are made either together or sequentially. For instance,
changes in the header file may also indicate that a change was made in the
implementation file. The visualization tool is designed to visualize implicit rules
about the order of operations on various files, if too many outliers are detected,
the project manager may want to restructure the project.  While this is useful for
new projects, the development model of the Linux kernel is already structured and is
unlikely to change without good reasons behind it.

Gource is a tool for providing an interactive timelapse of the state of a
repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
type of contribution a developer is making. These contribution types are one of,
adding a file, removing a file, and changing a file. While the timelapse is
interesting to watch, it does not provide any additional explanation of the changes
actually being made, only the frequency that they are being made and who is making
them.

The current industry standard tool for git repository visualization is Gitk. The
Gitk interface is built around the central DAG viewer. The DAG view displays the
commits on their respective branches, the author of the commit or merge, and the
date the commit or merge was authored. A user can select a single merge or commit
to view additional information. For both merges and commits, gitk will present the
user with the git log. If the user selected a commit, gitk will also present the
user with the patch information and the names of the files edited. gitk is unable to
provide the patches and filenames information in merges, as it is unable to
aggregate commit information in merges.

\section{Visualizing the merge-tree of Linux}

The goal is to build a tool that simplifies  navigation of the kernel commit information, while providing an explanation of the changes occurring at each commit
and, more importantly, indicates how commits are merged in their way to master.
We designed the tool with two use-cases in mind, though a user may switch
between the cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
These are users that are maintaining a section of the kernel and would like to pick
a merge (including all the commits that it merges) and merge it directly into their current repository. This is useful for
reducing the amount of re-implementation work. For these users, it is important to have the
ability to aggregate metadata about files and modules being effected by the merge.
\dmg{what is root?, what is leaves? perhaps remove unless you can be specific}
It is also important for these users to be able to navigate from the root toward the
leaves.

\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
These are users that start with a known merge or commit and would like to see what
other changes are being made in commits that are in the same merge. This is useful
to see what other commits are related to the current commit and how they get collated into merges that eventually end in master.
 This is primarily for maintainers that need to perform
some specific cherry picking of commits. We
must provide these users a mechanism for navigating from a single commit toward the
master, allowing them to see the bigger picture.

\subsection{Core Model}

In our visualizations, each commit is either made directly to Master, or is part of a Merge-tree, with its root in Master (a merge made by Linus Torvalds). Each merge may contain zero or more
children, which are other merges and commits.

To improve the visualization of the merge-tree, we transform the merge-tree into a new tree, where each commit points to its next merge in the way to Master. In
this model, commits have no children and are always leaf nodes. This allows the clustering of commits along their next merge, and avoid the space needed to
present long sequences of commits without merges in between.

Commits contain the
metadata for the changes made. This metadata includes the files changed, the lines
added and removed from each file, the author, the date the commit was merged into
the merge that led to being merged into the kernel, and the date the commit was
authored, the patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date. The details
of the model are outlined in \cite{German2015}.

\section{Design}

\dmg{I would remove the command line shell, and use the space for related work}

We came up with two designs. We recognized that the structure of the repository was
similar to the wide and shallow tree-structure of a file system. We began with a
command-line shell interface. Then transitioned to a web-based tool where we looked
beyond the file system structure. In this section, we describe the progression of
our tool design, showing the benefits and drawbacks of each component.

\subsection{Command-line Shell}

Our initial intuition was to work with the commits and merges as directory trees
where the commits are files and the merges are directories. There are many
similarities between the tree structure of a file system and the tree structure of
the repository. Both trees are generally shallow, but very wide. File systems have
been studied for quite some time and there are many tools capable of visualizing and
working with the file structure.

The first tool we implemented was a small command-line shell (figure
\ref{fig:shell}) that allows users to navigate through the commits and merges as one
might navigate through files and directories. A user could navigate toward the
leaves of the tree by using the \verb|cd| command followed by the commit ID of the
next merge. The user could then navigate toward the root by issuing \verb|cd ../|
as they would in a file system. Users can use \verb|ls| to view what commits and
merges exist in the current working merge, then use \verb|cat| to view the log
preview of a commit or \verb|more| to see the preview and how many lines were added
and removed from the files in the commit.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/shell.png}
        \caption{Shell Design}
        \label{fig:shell}
\end{figure}

The shell does not meet our design goals. It only provides users with the ability to
navigate from the root. If there is a commit or merge that they would like to
investigate, they must find it with another tool and navigate through the tree to
find it. Furthermore, it does not aggregate information in the merges, so users must
manually aggregate the metadata. Another issue with this design is that it presents
the users with the commit hashes, which are quite long and unpleasant to remember
and type into the shell, but also do not provide any quick explanation of what the
commit or merge may contain.

\subsection{Web-based Tool}

We replace the command-line interface with a web-based tool. Creating a web-based
tool enables users to use the system without having to install additional software
or store a large database, making it more accessible, more easily maintainable, and
platform independent. A web-based tool can utilize more interactive means of
navigation and provide better explanations of commits and merges than the
command-line interface.

Our tool uses the following mechanisms to reach our goals of better navigation and
better explanation of the changes.

\begin{itemize}
        \item Filter by search
        \item View files edited
        \item View modules
        \item Tree viewer
\end{itemize}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View}
        \label{fig:search}
\end{figure}

Searching (figure \ref{fig:search}) allows a user to filter commits and merges that
are irrelevant. The search mechanism breaks down the results by release version. A
user can further narrow down the search range using the commit date range. The
commit date refers to the date of the creation of the top-level merge. The commits
and merges returned may have a commit date beyond the queried dates; the date-range
query refers to the date that the commit is finally merged into the kernel by Linus.
A commit may not be merged immediately, and in some cases may take years before
being applied. A user may then provide a search text, filtering by the author name,
the commit ID, or keyword from the log. Any part of the author name may show up in
the results, including searching by email address. The commit ID must be specified
from the first character to the last character. For example, the commit
`c267548755a184ef97301071300c1739a564e135' is returned when the user searches for a
commit ID of `c26' in the 3.10 Linux kernel. We will use this commit for showing
each of the features of the tool.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results.png}
        \caption{Search Results}
        \label{fig:results}
\end{figure}

In the search results, the user is presented with the one-line log message preview,
the author's name and email, the date the commit was authored, and the date the
commit was last committed. The commit date and author date will usually be the same;
however, if the commit has been rebased or cherry-picked into a merge that
eventually is merged into the kernel, the commit date and authored date will be
different. We can see this in the first and third entries in figure
\ref{fig:results}.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/message_view.png}
        \caption{Commit Message}
        \label{fig:message}
\end{figure}

Once a user selects a commit or merge to investigate, they are presented with a
tabbed pane allowing them to view the full commit log, the files edited, the modules
involved, and the tree views.

The first tab displays the full commit log (figure \ref{fig:message}). From this, a
user is able to see what they would see had they searched for the commit using git
log. This doesn't provide additional information to the other tools, but helps to
complete the functionality of our tool. The commit log provides a user with the
information about the content of the commit and who has signed-off on the commit to
ensure that it is good quality. The message for merges may contain a summary of the
commits being merged.  The information within these messages is highly variable, and
is completely dependent on the author's style. As the user moves toward the
top-level merge, the quality of these messages generally improves.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/file_view.png}
        \caption{Commit Files}
        \label{fig:files}
\end{figure}

The second tab is the files tab (figure \ref{fig:files}). This tab provides
information on what files have been edited, how many lines were added, and how many
lines were removed in a given commit. This functionality is similar to the other
tools available. Our tree-based design model allows us to extend this functionality
to merges, which the other tools are unable to show. To find the number of lines
added to a file in a merge, we take the sum of the lines added to that file in each
of the children of that merge. We do the same for calculating the number of lines
removed. This could be extended to use the patch information to determine when a
line has been replaced rather than incrementing the lines added and removed.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules.png}
        \caption{Commit Modules}
        \label{fig:modules}
\end{figure}

The modules tab (figure \ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by git, and are not going to
be present in all repositories.  In the Linux repository, authors put the module
they are working on in the one-line preview of the log-message. We heuristically
extract the module by taking all text in the log preview of commits until the first
colon.  Modules are logical partitions of the information in the kernel. Depending
on where the author was working, modules can be general, such as ``bluetooth'' and
``wireless'', or can be quite specific for individual hardware, such as
``ath9k\_hw'' and ``wl1251''. In a few cases, the author of a commit does not
correctly follow this format and the hueristic approach fails. Like files, we are
able to show what modules are edited in merges and commits. The image in figure
\ref{fig:modules} shows the modules for a commit.  As commits have no children, they
will only have a single module; merges with more than one child will have more.

Finally, we have the tree view tab. The tree view is designed for providing easy
navigation within a tree and providing a clear topological view of the branch of the
main kernel. We have experimented with various tree designs to find a design that
allows for easy navigation and visualization of both large and small trees. We
discuss the various tree designs in section \ref{treeview_section}.

\subsection{Tree Views} \label{treeview_section}

The tree view is what makes this tool unique to other tools, providing a full view
of the entire merge tree. We have experimented with various types of trees; list
trees are similar to the representation used in many file system explorers,
Reingold-Tilford trees clearly represent the tree structure, and bubble trees
organize the data hierarchically by having the parent node contain the child nodes
similarly to tree maps, but clearly showing the parent-child relationships between
nodes.

\subsubsection{List Tree}

The list tree viewer (figure \ref{fig:list_tree}) is in the form of nested lists,
and is designed to more closely model the tree view found in file browsers. This
tree only contains the commits and merges that are within the subtrees of the
current merge. A commit will never have any items in this tree as it is a leaf. To
accompany the tree, we include breadcrumbs at the top of the page to enable a user
to navigate both from the root to the leaves and from a leaf to the root. The last
item in the breadcrumb list is the current commit, the previous item is the parent
of the current commit, and the first item is the root merge into the kernel.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
        \caption{List Tree}
        \label{fig:list_tree}
\end{figure}

\subsubsection{Reingold-Tilford Tree}

The Reingold-Tilford tree (figure \ref{fig:tree})produces very nice results,
creating an intuitive representation of the tree. This illustrates a clear notion of
root and leaves, and how to navigate in either direction.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
        \caption{Reingold-Tilford Tree}
        \label{fig:tree}
\end{figure}

Clicking on subtrees will cause them to minimize, only showing parts of the tree
that are of interest to the user. Furthermore, the user can zoom the tree by
scrolling the mouse wheel to see more of the topology of the tree, or to see the
details of a given part of the tree.

Some of the merge trees are very large (figure \ref{fig:zoomed_tree}), containing
thousands of commits and merges. While the tree is capable of producing a
visualization, it becomes much more difficult to understand.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
        \caption{Reingold-Tilford Tree zoomed out}
        \label{fig:zoomed_tree}
\end{figure}

\subsubsection{Bubble Tree}

The bubble tree (figure \ref{fig:bubble_tree}) provides a very clear picture of
where a commit is located in the hierarchy. The currently selected commit or merge
is colored red, while non-selected merges use a shade of blue and commits use white.
The depth of the blue colouring is dependent on how deep in the hierarchy that node
is. The root is the lightest blue, and the contained merges are progressively
darker.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
        \caption{Bubble Tree}
        \label{fig:bubble_tree}
\end{figure}

The bubble tree doesn't have an implicit way of providing additional information,
placing any text near the nodes makes the tree impossible to read, so we include a
separate pane in the web page. When a user hovers over a node, the pane shows
additional information about the author and the commit message and a link to the
detailed page for that commit or merge. If the user clicks on a node, the tree will
zoom to that node and the information in the info pane will persist, enabling the
user to click the link. This tree provides an easy mechanism for users to navigate
from the root to the leaves and vice versa.

\section{Implementation Details}

\evan{Is this section necessary? The tables are probably useful, but the rest... not
sure on}

We built a web-based tool to provide information about the kernel.  We chose to use
Nginx as the main web server, postgresql as the dbms, and Flask as the framework for
generating the dynamic content.

Nginx is a high-performance, open source, scalable HTTP and mail server. Started in
2002, the project is much younger than the Apache project, and designed with massive
scalability and multi-processing in mind. Instead of using request-driven threading,
Nginx uses an event-driven architecture, making it more scalable, proving itself as
being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.

Postgresql is our dbms of choice, providing fast and reliable results. There are
multiple reasons behind choosing postgresql over another dbms. The main reason being
that the dataset built by German et al. was generated from a postgresql database,
allowing us to import the data directly without any further conversions.

We chose the python micro-framework
Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
content. Working with flask is simple and easy, allowing us to work in Python.

The database itself is broken down into 5 main tables; commits, filesmod, logs,
pathtomerge, and releases. Commits contains most of the metadata for each commit; it
contains the author, the authored date, and committer, the commit date, a boolean
saying what is a merge, and the patch. Filesmod contains the information on what
files were modified for the commits. Since a file may be modified by multiple
commits, and a commit may be modifying multiple files, we can do one of two things
for the primary key. We could use both the filename and the commit ID as the primary
key, but we just use an index. The filesmod table contains the filename, the commit
ID, the number of lines added, and the number of lines removed. Trivially, it
contains the index, but that isn't necessary for anything beyond the primary key.
Logs contains the commit log preview and full commit log. Pathtomerge is the
important table.  Pathtomerge contains the commit ID, the number of merges between
the commit and the top-level merge with the kernel, the next merge in the merge path
up to the top-level merge with the kernel, the top-level merge into the kernel, and
the date it is merged into the kernel. The information in this table must be
generated periodically by the server because we are unable to generate the tree
structure correctly straight from the information stored by git. Pathtomerge does
not contain any top-level merges, as these will never have a next merge in the tree;
since we are only storing the parents and not the children, storing the top-level
merge is redundant and unnecessary. The releases table stores the release
information.

It contains
\begin{itemize}
        \item Version name
        \item Candidate releases versus real releases
        \item Previous release
        \item Previous real release
        \item Previous release commit ID
        \item Previous real release commit ID
        \item Current version commit ID
\end{itemize}
The current version commit ID represents the last top-level merge into the version
of the kernel. Using this and the commits table, we are able to determine the
release date of any version of the kernel. Then using the previous release date, we
can find the range of dates where valid top-level merges are found for that version.

\section{Discussion}

Our tool provides mechanisms and visualizations that other tools are unable to
produce. Continuous mining of the repository provides more information about how the
repository is changing. It gives us the ability to see how commits are being merged
into the kernel, allowing us to more reliably build trees from the snapshots of the
repository.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
        \caption{Merge Dag View}
        \label{fig:dag_view}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_tree.png}
        \caption{Merge Tree View}
        \label{fig:tree_view}
\end{figure}

Our goal was to build a tool to enable maintainers to effectively navigate and pull
an explanation of the changes performed to the kernel over the period of a release.
Achieving this goal includes removing information that does not pertain to the area
of the kernel that the user is interested in. In figure \ref{fig:dag_view} and
figure \ref{fig:tree_view} we can visually compare the results returned from gitk
and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
from kernel version 3.10. The relevant information to this commit in both of these
figures is identical, but the presentation drastically changes our ability to
comprehend what we are seeing. The primary difference is the removal of irrelevant
information. In gitk (\ref{fig:dag_view}), the visualization results include commits
and merges from other components of the kernel, while our tool (\ref{fig:tree_view})
only includes the results that are specific to the component of the kernel that we
are interested in.

This merge is relatively small, containing only four merges, three of which
only contain a single commit, and the last containing two commits. With our
tool, we are able to immediately see what commits and merges pertain to our
section of the kernel. The DAG view of this provides almost no explanation,
furthermore, users must work to determine where this merge ends and the next one
begins.

Our tool further enhances our understanding by immediately giving us the files
and modules that were edited in the branch. We are able to determine that three
files, ``palamas-regulator.c'' had two lines added and two lines removed,
``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
lines added and 2 lines removed. Finally, we are able to determine that only the
``regulator'' module was modified in this merge and was modified by 5 commits.

\subsection{Extendability}

The front-end uses asynchronous requests to gather the information for the trees and
tables. This enables third-parties to implement new front-ends, trying other
designs, though this interface may change in the future. The tree, file, and module
information is accessible through
http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.

The response for the tree is a single object. This object is the root node of the
tree, and may contain an object of children objects in the \verb|children| field,
which are in the same format as the root object. The children object uses the commit
ID as the key.

The tree node objects have the following fields:

\begin{tabular}{ccl}
        Field name      & Data type & Description\\\hline
        \verb|cid|      & string    & Commit ID\\
        \verb|name|     & string    & One-line log preview\\
        \verb|mlinus|   & string    & Root merge commit ID\\
        \verb|author|   & string    & Author name and email\\
        \verb|mnext|    & string    & Parent merge commit ID\\
        \verb|children| & object    & Object of tree objects\\
\end{tabular}

The file responses contain only the files that the selected merge or commit works
with. The response is a single object in the form of a tree. This object is the root
of the tree, and represents the current merge or commit. If the current position in
the tree is an inner node, the response will contain the child nodes in the
``children'' field, otherwise ``children'' will be an empty list. The format of file
response is as follows,

\begin{tabular}{lll}
        Field name & Data type & Description\\\hline
        \verb|cid| & string & Commit ID\\
        \verb|mnext| & string & Parent merge commit ID\\
        \verb|children| & list & List of tree objects\\
        \verb|files| & list of tuples&
        \footnotesize{
        \begin{tabular}{lc}
                \verb|Filename| & string\\
                \verb|Lines added| & uint\\
                \verb|Lines removed| & uint\\
        \end{tabular}}\\
\end{tabular}

\section{Future Work}

There are still many areas that can be improved before the full potential of our
model can be realized. Here, we outline various areas of the tool that still need
more attention.

\subsection{Files}

There currently is no functionality surrounding searching by filename. Users falling
under use-case 2 may know what file they are editing and try to determine how it may
work with the other commits and merges in the module. It is possible that a third
use-case may arise, where the user wants to determine all commits that effect a
given file. In both cases, bottom-to-top approach is applied.

There is limited functionality to the presentation of the file information. At a
minimum, the patches for the commits can be displayed. From there, the patches can
be used to piece together parts of the file to generate a single patch at a given
merge rather than displaying each patch individually. The patches can also be used
for determining what kinds of changes were made, if the lines are being added,
removed, or being replaced.

\subsection{Authorship}

Our model can aggregate more information than what we have implemented. The
authorship information is important for licensing purposes, and we can show all
authors contributing to commits in a  merge and how many commits they contributed to
the merge. We could go further, providing information about what files they edited,
how many lines they added and removed, and more.

\subsection{Evaluation}

At this time, we have no evidence that our tool is able to improve the work-flow of
maintainers. We believe that the tool is able to improve the work-flow and
performance of maintainers because it provides cleaner mechanisms of visualizing and
presenting the commit information. It is able to provide more relevant information,
while removing information that is irrelevant to a given module or set of merges.

We can either perform user-testing to show how a users workflow changes, or we could
have maintainers evaluate and critique the tool and use their feedback to determine
if the tool meets our goals.

\section{Conclusion}

Our tool shows promising results in working toward our goals, building a tool that
can more easily navigate the repository, and provide clearer explanations of the
changes made. The filtering by searches and the various tree visualizations enable
users to easily navigate through the kernel, finding the commits and merges that
pertain to what they are interested in. The tree visualizations are able to both
serve as a navigation piece, providing simpler navigation through the kernel, but
also demonstrate which commits contribute to which merges, providing further
explanation over the DAG. The tree model further improves the explanation by enabling
the tool to aggregate metadata at the merges instead of requiring a user to manually
aggregate the information. Our tool currently aggregates file information and the
modules edited, but can be extended to include authorship information and aggregated
patch information.

The tree model requires more information than what is provided by git, requiring a
central server to monitor changes to the project. This breaks the design goals of
decentralized version control systems and makes the visualization tool specific to
the project.

While these are significant drawbacks for small projects, large projects will likely
have a more centralized work flow with a master repository where contributions will
eventually land, along with various servers that can watch this master repository
for changes. Given that the importance of clear visualizations of the repository
outweighs the cost of having a server monitor the state of the repository, the
tree-based model appears to provide clean visualizations and explanations of the
changes in the Linux kernel.

\nocite{*}

\bibliographystyle{IEEEtran}
\bibliography{citations}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: