### 2016-04-26 linvis.tex
Minor capitalization changes

% vim: set fenc=utf-8 ft=latex encoding=utf-8
% -*- mode: latex; coding: UTF-8; -*-

\newif\ifdraft
% \drafttrue

\ifdraft
        \documentclass[conference, draftclsnofoot, draft]{IEEEtran}
        \def\baselinestretch{1}
        \setlength{\marginparwidth}{2cm}
\else
        \documentclass[conference]{IEEEtran}
\fi

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

\newcommand{\TheTitle}{Merge-tree: Visualizing the integration of commits into Linux}
\newcommand{\TheAuthors}{Evan Wilde, Daniel German}
\newcommand{\TheEmails}{etcwilde@uvic.ca, dmg@uvic.ca}
\newcommand{\TheSubject}{Digesting large amounts of commit data}
\newcommand{\TheKeywords}{Linux, git, data structures, tree data structures}

\synctex=1

\usepackage[hyphens]{url}
\urlstyle{same}

\ifdraft
        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
                pdfborder={0 0 0},backref=none,colorlinks=true]{hyperref}

\else
        \usepackage[unicode=true,bookmarks=false,breaklinks=false,
                pdfborder={0 0 0},backref=none,colorlinks=false]{hyperref}
\fi

\usepackage[nospace]{cite}

% Table Support
\usepackage{dcolumn}
\usepackage{longtable}

\usepackage{balance}
\usepackage{placeins}
\usepackage{multirow}

% Extra support
\usepackage{xspace}
\usepackage{caption}
\usepackage[nospace]{cite}

\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}

% Fix any bad-hyphenations here
\hyphenation{contained}

\input{macros.tex}

\begin{document}

\title{\TheTitle}
\author{
        \IEEEauthorblockA{\TheAuthors}
        \IEEEauthorblockN{Department of Computer Science,
                University of Victoria, Canada.}
        \IEEEauthorblockA{Email: \TheEmails}
}
\maketitle
\begin{abstract}

        With an average of more than 900 top-level merges into the Linux kernel per
        release, many containing hundreds of commits and some containing thousands,
        maintenance of older versions of the kernel becomes nearly impossible.
        Various commercial products, such as the Android platform, run older
        versions of the kernel. Due to security, performance, and changing hardware
        needs, maintainers must understand what changes (commits) are added to the
        current version of the kernel since the last time they inspected it in order
        to make the necessary patches.

        Current tools provide information about repositories through the directed
        acyclic graph (DAG) of the repository, which is helpful for smaller
        projects. However, with the scale and number of branches in the kernel the
        DAG becomes overwhelming very quickly. Furthermore, the DAG contains every
        ancestor of every commit, while maintainers are more interested in how and
        when a commit arrives to the official Linux repository.

        In this paper, we propose the merge-tree view, a simplified transformation
        of the DAG of the Linux git repository that shows the way in which commits
        are merged into the master branch of Linux. Using the merge-tree view we
        build \tool, a tool that is designed to allow users to explore how commits
        are merged into the Linux kernel.

\end{abstract}

\begin{IEEEkeywords}
        \TheKeywords
\end{IEEEkeywords}

\section{Introduction}

\evantodo {
        \begin{enumerate}
                \item Check captions
        \end{enumerate}
}

Between 50k and 70k commits are added to the Linux kernel per version
requiring maintainers of older versions of the Linux kernel to sift through
thousands of commits and merges with tools that are unable to filter and effectively
visualize projects at the scale of the kernel. Older versions of the kernel are used
in embedded systems and mobile phones; for security purposes, performance needs, and
changing hardware requirements, maintainers must be able to understand the changes
being made in the current version of the kernel in order to produce the necessary
patches for the older versions of the kernel. Tools like Gitk use a directed acyclic
graph (DAG) model of the repository, showing all commits and merges in chronological
order by when the commit was authored, not by when it arrived in the official Linux
repository.

This representation works in smaller projects; it enables users to see when changes
are made, when these changes are merged, how each branch is interacting, and the
point where a branch forks from the master branch. In large modular projects, like
the Linux kernel, the DAG becomes a mess of merges and commits
(Figure~\ref{fig:gitk}) losing its visual meaning. In some cases, the Linux kernel
is simply too large for the system to generate a visualization; Github provides a
DAG view for many projects, but is unable to display the visualization for projects
at the scale of the Linux of the kernel (Figure~\ref{fig:gitfail}).  Between 60 and
70k new commits are created for the Linux project every year; According to our
previous work\cite{German2015}, a commit takes a median of 30 days from the time it
is authored until it arrives in the official repository. The snapshot of the kernel
tomorrow may be different than the snapshot from today, containing new commits
authored in the past; distinguishing these new commits from the commits in the
snapshot from today is not trivial.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/gitk.png}
        \caption{A sample DAG view of commits from Linux in Gitk. Commits point
                towards their ancestors, and there is no clear indication of the way
                the travel to reach master. In fact, neither Gitk nor Git is capable
                of showing what commits were done in master.}
        \label{fig:gitk}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.45\textwidth]{figures/github_viewer.png}
        \caption{Github Failure Message showing the DAG of the Linux Project and its
                relationship to other forks.}
        \label{fig:gitfail}
\end{figure}

One major challenge with visualizing the arrival of commits to a repository is that
Git does not store the date that a commit was merged into another branch (including
master). To complicate the problem, the DAG has reference to the ancestors of a
commit (a model necessary for the operation of Git), but developers would prefer
knowing the path a commit followed to reach the master repository. Tracing a path
that any commit followed to the master repository would imply that for any given
merge, it would be possible to know which commits were merged. A user could inspect
the commits that arrived into master within a given time-frame by checking which
commits were merged during that time-frame.

This paper makes two contributions; first, we describe a method of converting the
DAG of the Linux repository into a tree, or \emph{merge-tree} of the repository,
that represents the path used by a commit to reach master; second, we present a
method to inspect and visualize the history of merges in the Linux project using the
merge-tree model.

These methods and visualizations are implemented in a web-based tool called
\tool\footnote{\tool is currently available at \url{http://li.turingmachine.org}}.
Our visualizations and tool provide information about the location of any given
commit in its respective merge-tree, the files edited, the modules edited, and the
commit message for a given commit or merge. \tool allows users to apply various
filters, including the release version, along with a keyword or phrase from the log
preview, the name of the author, or the commit ID. The user can request all the
root-level merges containing a commit or merge that matches the query, or all
commits and merges that match the query.

\section{Merge-Tree model}
\label{sec:mergetree}

Git uses a directed acyclic graph (DAG) as its main data model. In this model, a
commit has one or more ordered parents (ancestors), except the root commits of a
repository which do not have ancestors--Linux git repository has two such commits.
Commits are divided into merge commits, merging two or more branches, and non-merge
commits. Non-merge commits have only one parent, while merge commits have two or
more. The order of the parents matter: the first parent is the branch in which the
merge is being done, while the rest indicate the branches being merged (Git allows
merging multiple branches simultaneously). The second parent is the first branch
merged, the third parent is the second branched merged.

Once a commit is created, it is never changed. Git allows operations to alter
commits or reorder them, but it changes the commit ID in the process, effectively
making a new commit. This makes commits unable to record the traversal of merges
from the commit to the merge into the master branch.

A short example: assume the of commits represented in Figure~\ref{fig:repoEvents}
shows the sequence of events in a repository. In this case, commits are performed in
various repositories and branches. The DAG representation of the commits is shown in
Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about the master
branch and the repository that the master branch is part of. The merge-tree view of
this DAG is visible in Figure~\ref{fig:repoTree}. Note that the direction of the
edges of the DAG has been inverted, instead of pointing from the child to the
ancestors, it points from the ancestor to its successors, forming a path to the
master branch. Also note that the DAG has been simplified, showing only a single
edge on the path to master for any commit.

\begin{figure}[htbp]
        \centering
        \includegraphics[width=\columnwidth]{figures/events.pdf}
        \caption{Example of a sequence of events performed in different
                repositories. The horizontal axis represents time. Each horizontal
                section represents a different branch and/or repository. Each commit
                points to its ancestor.}
        \label{fig:repoEvents}
\end{figure}

\begin{figure}[htbp]
        \centering
        \includegraphics[width=.8\columnwidth]{figures/dag.pdf}
        \caption{DAG representation of the commits represented in
                Figure~\ref{fig:repoEvents}. It loses information about what
                repository the commit is performed and through which repositories it
                has passed in its way to master. In fact, the DAG does not even know
                what master is.}
        \label{fig:repoDAG}
\end{figure}

\begin{figure}[htbp]
        \centering
        \includegraphics[width=\columnwidth]{figures/tree.pdf}
        \caption{Merge-tree view of the commits  represented in
                Figure~\ref{fig:repoEvents} showing the path they followed to reach
                master. In this model the successors of each commit represents the
                path followed by that commit to reach master.}
        \label{fig:repoTree}
\end{figure}

\subsection{Computing the merge-tree of the DAG of Linux}

Computing the merge-tree from a DAG for any repository may not be possible; however,
certain features of the development process of Linux make it feasible to compute the
merge-tree for the Linux repository. First, the master branch of Linux is maintained
by Linus Torvalds, and only Linus has write access to it. We have verified this
assertion in previous research~\cite{German2015}. We have developed a heuristic that
is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first identifies
the commits made directly to the master branch; whereafter it recursively determines
the shortest path (in terms of time), using the DAG, from each commit to master
using the inverted DAG.

\begin{algorithm}
        \caption{Computing the merge-tree of Linux Git's DAG}\label{fig:alg}
        \begin{algorithmic}
                \Function{ComputeMergeTree}{DAG}: tree
                \State {\# Compute the tree from the DAG of Linux repository.}
                \State {\# Returns $Tree$, a graph containing every commit }
                \State {\# in DAG with the path it followed to master.\\}

                \State $head \gets \textit{Head of master of git repository}$
                \State $master \gets \textit{traverse DAG from head using }$
                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
                \State $nodes(Tree) \gets nodes(DAG)$
                \State \Function{distance2Master}{cid} : seconds
                \State {\# Helper function}
                \State {\# Recursively compute shortest distance to master}
                \State {\# setting cid's successor (next) in its way to master.}
                \State {\# This function should be memoized. Otherwise it}
                \State {\# would run in exponential time.}
                \If {\textit{cid in master}}
                \State \Return 0
                \EndIf
                \State    $d \gets 	\infty$
                \State {\# Traverse the inverted DAG}
                \For{$c \in children(cid, DAG)$}
                \If {$c \in master$}
                \State $d_1 \gets commitTime(c)-commitTime(cid)$
                \Else
                \State {$d_1 \gets distance2Master(c)$}
                \EndIf
                \If {$d_1 < d $}
                \State $next \gets c$
                \State  $d \gets d_1$
                \EndIf
                \EndFor
                \State {\# $c$ is the commit that follows $cid$}
                \State {\# in its way to master}
                \State add edge $(cid, next)$ to $Tree$
                \State \Return $d$
                \EndFunction

                \State {\# Compute the distance for each commit}
                \State {\# discarding result}
                \For{$c \in nodes(DAG)$}
                \State $distance2Master(c)$
                \EndFor
                \State \Return $Tree$
                \EndFunction
        \end{algorithmic}
\end{algorithm}

\subsection{Evaluation}

Merges that do not have conflicts provide information to verify this heuristic. If a
merge does not contain a conflict, it records a summary of the commits that it
merges. See Figure~\ref{fig:sampleMerge} for an example. This summary contains a
list of the first 20 non-merge commits in the merge, including their one-line log
description, the full logs of the merge commits that merge this subset, and the
total number of non-merge commits in the merge.

\begin{figure}[htbp]
        \centering
        {\tiny
        \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

Merge branch 'for-linus' of git://neil.brown.name/md

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache when mddev has no gendisk
...
        \end{verbatim}}
        \caption{Example of how merges record a subset of commits being merged. The
                commit only shows the first 20 one line log messages for the 24
                non-merge commits it merged. The ending ... is part of the log and
                represents that other commits were merged.}
        \label{fig:sampleMerge}
\end{figure}

We use this information to evaluate the accuracy of the merge-tree model extracted
from the DAG. For commits with no conflicts, we compare: 1) the number of non-merge
commits, 2) the summaries of the commits available in the log, and 3) the summaries
of the merges included.

The verification detected no incorrect merges since 2013, two incorrect merges in
2012, and two in 2011; however, the heuristic breaks in October 15 2009 due to the
presence of \textit{foxtrot} commits, which confounds the true master branch of a
repository
\footnote{See \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html} for a description of the issue}, and will not work for older merges.

\section{Visualizing the merge-tree of Linux}

The goal of \tool is to simplify the navigation of the kernel commit information,
specifically focusing on merges. This is done by leveraging the merge-tree view to
inspect how commits are merged on the path to the master repository.

\subsection{Use cases}

We designed \tool with two use-cases in mind, though a user may switch between the
cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom approach}\label{sec:usecase1}\\
These are users that are maintaining a section of the kernel and would like to pick
a merge (including all the commits that it merges) and merge it directly into their
current repository. This is useful for reducing the amount of re-implementation
work. For these users, it is important to have the ability to aggregate metadata
about files and modules being effected by the merge.  It is also important for these
users to be able to navigate from the root of the merge-tree toward the leaves.

\noindent \textbf{Use-case 2: bottom-to-top approach}\label{sec:usecase2}\\
These are users that start with a known merge or commit and would like to see what
other changes are being made in commits that are in the same merge, including
knowing the merge-tree they belong to. This is useful to see what other commits are
related to the current commit and how they get collated into merges that eventually
end in master.  This is primarily for maintainers that need to perform some specific
cherry picking of commits. We must provide these users a mechanism for navigating
from a single commit toward the master, allowing them to see other commits that
might be related to it.

\subsection{Data Model}

In our visualizations, we leverage the Merge-Tree model described in
Section~\ref{sec:mergetree}. In this model each commit is either already in master,
or is part of a tree which is rooted in the merge that merged it into master.  Each
commit, whether a merge or non-merge, has only one successor; the root of each tree
has none as it was made by Linus Torvalds directly into the master branch. Non-merge
Commits contain the metadata for the changes made. This metadata includes the files
changed, the lines added and removed from each file, the author, the date the commit
was merged into the merge that led to being merged into the kernel, the date the
commit was authored, the patch, and the commit log. Merges contain less metadata,
only storing the author of the merge, the log, the commit date, and the author date,
and potentially, changes necessary to address conflicts during the merge. The
details of the model are outlined in \cite{German2015}.

\section{Design and Implementation}

To navigate and inspect the merge-tree view of the Kernel we created a web-based
tool called \tool. Creating a web-based tool enables users to use the system without
having to install additional software or store a large database, making it more
accessible, more easily maintainable, and platform independent. \tool uses the
following mechanisms to reach our goals of better navigation and better explanation
of the changes.

\begin{itemize}
        \item Filter by searching
        \item View files edited
        \item View modules
        \item Tree viewer
\end{itemize}

\subsection{Searching}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search.png}
        \caption{Search View. It allows finding commits that were \textbf{merged} in
                a given period or by commit it.}
        \label{fig:search}
\end{figure}

Searching (depicted in Figure \ref{fig:search}) allows a user to filter commits and
merges that are irrelevant. The search mechanism breaks down the results by release
version. A user can further narrow down the search by \textbf{specifying a range of
dates} in which such commits \textbf{were merged by Linus into master, not when the
commits were created.} This distinction is important. We have observed commits that
have taken years to arrive into master after they were originally committed.

A user may then provide a search text, filtering by the author name, the commit ID,
or keyword from the log. Any part of the author name may show up in the results,
including searching by email address.

If providing commit ID must be specified from the first character to the last
character. For example, the commit `c267548755a184ef97301071300c1739a564e135' is
returned when the user searches for a commit ID of `c26' in the 3.10 Linux kernel.
This commit is a non-merge commit performed, in one of the largest merge-trees of
the kernel; however, this feature will work with any commit ID.  We will use this
commit for showing each of the features of \tool.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/search_results_2.png}
        \caption{Search Results. Each is a commit merged in the desired merged
                window.}
        \label{fig:results}
\end{figure}

In the search results, the user is presented with the one-line log message preview,
the author's name and email, the date the commit was authored, and the date the
commit was last committed. The commit date and author date will usually be the same;
however, if the commit has been rebased or cherry-picked into a merge that
eventually is merged into the kernel, the commit date and authored date will be
different. We can see this in the first entry in Figure~\ref{fig:results}.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/log_view.png}
        \caption{Panel showing the Commit Message of a commit.}
        \label{fig:message}
\end{figure}

Once a user selects a commit or merge to investigate, they are presented with a
tabbed pane allowing them to view the full commit log, the files edited, the modules
involved, and the merge-tree view.

The first tab displays the full commit log (Figure~\ref{fig:message}). From this, a
user is able to see what they would see had they searched for the commit using Git
log. This doesn't provide additional information to the other tools, but helps to
complete the functionality of \tool. The commit log provides a user with the
information about the content of the commit and who has signed-off on the commit to
ensure that it is good quality. The message for merges may contain a summary of the
commits being merged.  The information within these messages is highly variable, and
is completely dependent on the author's style. As the user moves toward the
top-level merge, the quality of these messages generally improves.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/files_view_2.png}
        \caption{Panels showing the files modified by all the commits that are part
                of this merge.}
        \label{fig:files}
\end{figure}

The second tab is the files tab (Figure~\ref{fig:files}). This tab provides
information on what files have been edited, how many lines were added, and how many
lines were removed in a given commit. For non-merges, this functionality is similar
to the other tools available. Our tree-based design model allows us to extend this
functionality to merges by aggregating information about all the commits that are
children of the merge in the merge-tree, which other tools are unable to show. To
find the number of lines added to a file in a merge, we take the sum of the lines
added to that file in each of the children of that merge. We do the same for
calculating the number of lines removed.

% This could be extended to use the patch information to determine when a line has been
% replaced rather than incrementing the lines added and removed.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/modules_view_2.png}
        \caption{Panel showing the modules changed by all the commits in this merge.}
        \label{fig:modules}
\end{figure}

The modules tab (Figure~\ref{fig:modules}) shows the modules that are contained
within the commit. Modules are not natively recognized by Git, and are not going to
be present in all repositories. In the Linux repository, authors put the module they
are working on in the one-line preview of the log-message; for example: the log
\emph{gcov: add support for FCC 4.9} has updated the \emph{gcov} module (the
coverage testing tool of the kernel). We heuristically extract the module by taking
all text in the log preview of commits until the first colon.  Modules are logical
partitions of the information in the kernel. Depending on where the author was
working, modules can be general, such as ``bluetooth'' and ``wireless'', or can be
quite specific for individual hardware, such as ``ath9k\_hw'' and ``wl1251''. In a
few cases, the author of a commit does not correctly follow this format and the
heuristic approach fails. As with the Files panels, non-merge commits show their
corresponding information, but for merge commits we aggregate all the modules
changed in all the commits that are part of that merge.  The output of this view is
shown in Figure~\ref{fig:modules}.

Finally, we have the Tree view tab. The tree view is designed for providing easy
navigation of the commits within the merge-tree that is rooted in the current merge.
It also provides a clear topological view of the merge and the submerges it
includes. We have experimented with various tree designs to find a design that
allows for easy navigation and visualization of both large and small trees. We
discuss this panel in the next subsection.

\subsection{Merge-Tree Views} \label{treeview_section}

The merge-tree view is what makes \tool unique to other tools that inspect the DAG
of a Git repository. With it, a user can inspect how commits are merged on their way
to the master branch. We have experimented with various types of trees:

\begin{enumerate}
        \item List trees are a text-based representation of the merge-tree, and are easy to search and navigate.
        \item Reingold-Tilford trees provide a clear visual represent the tree structure of the merge-tree.
        \item Bubble trees organize the data hierarchically by having the parent node contain the child nodes similarly to tree maps, but
                clearly showing the parent-child relationships between commits and merges.
\end{enumerate}

\subsubsection{List Tree}

The list tree viewer (Figure~\ref{fig:list_tree}) is in the form of nested lists,
and is designed to more closely model the tree view found in file browsers. This
tree only contains the commits and merges that are within the subtrees of the
current merge. A commit will never have any items in this tree as it is a leaf. To
accompany the tree, we include breadcrumbs at the top of the page to enable a user
to navigate both from the root to the leaves and from a leaf to the root. The last
item in the breadcrumb list is the current commit, the previous item is the parent
of the current commit, and the first item is the root merge into the kernel.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/list_tree.png}
        \caption{List Tree}
        \label{fig:list_tree}
\end{figure}

\subsubsection{Reingold-Tilford Tree}

The Reingold-Tilford tree\cite{Reingold1981} (Figures \ref{fig:zoomed_tree} and
\ref{fig:tree}) allows the visualization and navigation of the entire merge-tree in
an intuitive representation of the tree.  This illustrates a clear notion of root
and leaves, and how to navigate in either direction. Some of the merge trees are
very large (Figure~\ref{fig:zoomed_tree}), containing thousands of commits and
merges. While the tree is capable of producing a visualization, it becomes much more
difficult to understand. For example the merge
73287a43cc79ca06629a88d1a199cd283f42456a performed by Linus Torvalds in May 1, 2013,
merged 1722 non-merge commits and 91 merges. \footnote{
        This commit can be inspected at
        \url{http://li.turingmachine.org/commits/73287a43cc79ca06629a88d1a199cd283f42456a}}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_zoom.png}
        \caption{Entire Reingold-Tilford Tree of a merge. This merge by Linus (73287a43cc79ca06629a88d1a199cd283f42456a) contains 1722 includes non-merge commits
                touching 239 modules and 1525 files.}
        \label{fig:zoomed_tree}
\end{figure}

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/tree_view.png}
        \caption{Detail of the Reingold-Tilford Tree in Figure~\ref{fig:zoomed_tree}.}
        \label{fig:tree}
\end{figure}

The tree can be further extended to include the collapsing of merge nodes. Clicking
on merge nodes will cause them to collapse, only showing parts of the tree that are
of interest to the user.Furthermore, the user can zoom the tree by scrolling the
mouse wheel to see more of the topology of the tree, or to see the details of a
given part of the tree.

\subsubsection{Bubble Tree}

Bubble trees are useful for providing a clear visualization of hierarchal
data\cite{Boardman2000}. The bubble tree (Figure~\ref{fig:bubble_tree}) provides a
very clear picture of where a commit is located in the tree hierarchy. The currently
selected commit or merge is colored red, while non-selected merges use a shade of
blue and commits use white.  The depth of the blue colouring is dependent on how
deep in the hierarchy that node is. The root is the lightest blue, and the contained
merges are progressively darker.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/bubble_tree.pdf}
        \caption{Bubble Tree of a large merge, showing in red (the tiny dot close to the middle) the location of a specific commit of interest.}
        \label{fig:bubble_tree}
\end{figure}

The bubble tree doesn't have an implicit way of providing additional information,
placing any text near the nodes makes the tree impossible to read, so we include a
separate pane in the web page. When a user hovers over a node, the pane shows
additional information about the author and the commit message and a link to the
detailed page for that commit or merge. If the user clicks on a node, the tree will
zoom to that node and the information in the info pane will persist, enabling the
user to click the link. This tree provides an easy mechanism for users to navigate
from the root to the leaves and vice versa.

\section{Implementation Details}

\tool was built using Nginx as the main web server, postgresql as the DBMS, and
Flask as the framework for generating the dynamic content.

Nginx is a high-performance, open source, scalable HTTP and mail server. It is designed with massive
scalability and multi-processing in mind. Instead of using request-driven threading,
Nginx uses an event-driven architecture, making it more scalable, proving itself as
being the web server behind large sites like Netflix, Hulu, GitHub, Tumblr, and many
other popular sites\footnote{\url{http://wiki.nginx.org/Main}}.

Postgresql is our DBMS of choice, providing fast results.

We chose the python micro-framework
Flask\footnote{\url{http://flask.pocoo.org/docs/0.10/}} for generating the dynamic
content. Working with flask is simple and easy, allowing us to work in Python.

\subsection{Extendability}

The front-end uses asynchronous requests to gather the information for the trees and
tables. This enables third-parties to implement new front-ends, trying other
designs, though this interface may change in the future. The tree, file, and module
information is accessible through
http://li.turingmachine.org/data/tree/JSON/\verb|<cid>|,
http://li.turingmachine.org/data/files/JSON/\verb|<cid>|,
http://li.turingmachine.org/data/modules/JSON/\verb|<cid>|, respectively.

The response for the tree is a single object. This object is the root node of the
tree, and may contain an object of children objects in the \verb|children| field,
which are in the same format as the root object. The children object uses the commit
ID as the key.

The tree node objects have the following fields:

\begin{tabular}{ccl}
        Field name      & Data type & Description\\\hline
        \verb|cid|      & string    & Commit ID\\
        \verb|name|     & string    & One-line log preview\\
        \verb|mlinus|   & string    & Root merge commit ID\\
        \verb|author|   & string    & Author name and email\\
        \verb|mnext|    & string    & Parent merge commit ID\\
        \verb|children| & object    & Object of tree objects\\
\end{tabular}

The file responses contain only the files that the selected merge or commit works
with. The response is a single object in the form of a tree. This object is the root
of the tree, and represents the current merge or commit. If the current position in
the tree is an inner node, the response will contain the child nodes in the
``children'' field, otherwise ``children'' will be an empty list. The format of file
response is as follows,

\begin{tabular}{lll}
        Field name & Data type & Description\\\hline
        \verb|cid| & string & Commit ID\\
        \verb|mnext| & string & Parent merge commit ID\\
        \verb|children| & list & List of tree objects\\
        \verb|files| & list of tuples&
        \footnotesize{
                \begin{tabular}{lc}
                        \verb|Filename| & string\\
                        \verb|Lines added| & uint\\
                        \verb|Lines removed| & uint\\
                \end{tabular}}\\
\end{tabular}

\section{Discussion}

From a developers point of view, the are two major disadvantages of the DAG model of
Git: a) its edges point backwards, i.e. commits point to their ancestor, not to the
commit that succeeds them on the path to being integrated; and b) the DAG contains
many more edges that are necessary to understand how integration occurred. We
addressed these two issues with the creation of the merge-tree model of the DAG of
Linux. Effectively the merge-tree view recovers the details of each merged branch
into master.

In our experience, no other Git repository reaches the level of DAG complexity that
Linux has. In most Git repositories, merges are not nested, all merges merge
directly into the master repository, and these merges only merge few commits. Even
in these simplified cases, the merge-tree can provide a valuable summary of how
commits are integrated into the master branch, specifically since the time of
integration may be very different from the time the commits were authored.

The biggest challenge, however, is computing the merge-tree. Our heuristic
\textbf{only} applies to the Linux kernel repository, and only since July 2009. It
is very likely that the heuristic will not work with other repository, specially
those with fast-forward commits into the master branch (a majority of repositories).
We can address the lack of information by monitoring the state of the repository at
short intervals as we did in \cite{German2015}. This continuous mining of the
repository requires foresight and planning.

Because \tool leverages the merge-tree model of the kernel, it provides mechanisms
and visualizations that other tools are unable to produce.It gives us the ability to
see how commits are being merged into the kernel, allowing us to better understand
how code is integrated into the kernel.

\begin{figure}
        \centering
        \includegraphics[width=0.47\textwidth]{figures/042dd_DAG.png}
        \caption{Merge Dag View}
        \label{fig:dag_view}
\end{figure}

\begin{figure*}
        \centering
        \includegraphics[width=0.97\textwidth]{figures/042dd_tree.png}
        \caption{Merge Tree View \dmg{this Figure is unreadable. Any chance you can make the text bigger?}}
        \label{fig:tree_view}
\end{figure*}

Our goal was to build a tool to enable maintainers to effectively navigate and
browse the changes performed to the kernel over the period of a release.  Achieving
this goal includes removing information that does not pertain to the area of the
kernel that the user is interested in. In Figure~\ref{fig:dag_view} and
Figure~\ref{fig:tree_view} we can visually compare the results returned from Gitk
and our tool for the top-level merge ``042dd60ca6dec9a02cefa8edd67de386e35755d6''
from kernel version 3.10. Note that Gitk has no way to show or recognize that this
is a merge into the master branch.

The log information to this merge-commit in both of these figures is identical, but
the presentation drastically changes our ability to comprehend what we are seeing.
The primary difference is the removal of irrelevant information. In Gitk
(Figure~\ref{fig:dag_view}), the visualization results include commits and merges
from other components of the kernel, while our tool (Figure~\ref{fig:tree_view})
only includes the results that are specific to the component of the kernel that we
are interested in.

This merge is relatively small, containing only four sub-merges, three of which
only contain a single commit, and the last containing two commits. With our
tool, we are able to immediately see what section of the kernel these commits and merges pertain to. The DAG view of this provides almost no explanation,
furthermore, users must work to determine where this merge ends and the next one
begins.

\tool further enhances our understanding by summarizing the files
and modules that were edited in the entire branch being merged. We are able to determine that three
files, ``palamas-regulator.c'' had two lines added and two lines removed,
``dbx500-prcmu.c'' had 12 lines added and 12 lines removed, and ``core.c'' had 5
lines added and 2 lines removed. Finally, we are able to determine that only the
``regulator'' module was modified in this merge and was modified by 5 commits.

The merge-tree model used in \tool has many advantages to the DAG model, but there
are two drawbacks. The first drawback is that a server must query the state of the
master repository since time information is not stored by Git in order to properly
construct the merge tree. The second drawback is the need to pre-compute the
merge-tree of the repository.

% Our core model provides advantages to the DAG model used by other tools. We separate
% the commits and merges into sub-trees and focus on a single sub-tree at a time. This
% enables us to filter out commits and merges that do not pertain to the component of
% the kernel that we are interested in, providing a clearer picture of what is
% happening in a given merge. The tree model also enables us to aggregate metadata
% easily. With the addition of the continuous polling of the repository, the model
% provides us with a mechanism of performing date-range queries with meaningful
% results.

\section{Future Work}

There are still many areas that can be improved before the full potential of our
model can be realized. Here, we outline various areas of the tool that still need
more attention.

\subsection{Files}

There currently is no functionality surrounding searching by filename. Users falling
under use-case 2 may know what file they are editing and try to determine how it may
work with the other commits and merges in the module. It is possible that a third
use-case may arise, where the user wants to determine all commits that effect a
given file. In both cases, bottom-to-top approach is applied.

There is limited functionality to the presentation of the file information. At a
minimum, the patches for the commits can be displayed. From there, the patches can
be used to piece together parts of the file to generate a single patch at a given
merge rather than displaying each patch individually. The patches can also be used
for determining what kinds of changes were made, if the lines are being added,
removed, or being replaced.

\subsection{Authorship}

Our model can aggregate more information than what we have implemented. The
authorship information is important for licensing purposes, and we can show all
authors contributing to commits in a  merge and how many commits they contributed to
the merge. We could go further, providing information about what files they edited,
how many lines they added and removed, and more.

\subsection{Evaluation}

At this time, we have no evidence that our tool is able to improve the work-flow of
maintainers. We believe that the tool is able to improve the work-flow and
performance of maintainers because it provides cleaner mechanisms of visualizing and
presenting the commit information. It is able to provide more relevant information,
while removing information that is irrelevant to a given module or set of merges.

We can either perform user-testing to show how a users workflow changes, or we could
have maintainers evaluate and critique the tool and use their feedback to determine
if the tool meets our goals.

\section{Related Work}

Version control systems monitor the development lifetime of software projects. This
makes the version control system vital in providing information about how a software
project is being developed. To our knowledge, we know of no Git repository
visualization tool that builds a tree that maps the path a commit follows to the
master branch from the DAG provided by Git. This may be because more information is
required to generate the merge-tree model than what is stored by Git. However, there
has been a lot of work in providing visualizations of various repositories.

Many tools work to address the issues in communication between developers in
inter-team collaboration work. Codebook \cite{Begel2010} uses a data mining
technique  to determine the developer of a piece of code, the program manager who
wrote the specification for the code, and the program managers and developers on the
team who were working together. Hipikat\cite{Cubranic2005} is another tool with a
focus on communication. Where codebook focuses on developers working on a project,
Hipikat is focused on enabling easier integration of new developers to a project by
providing them with easily-searchable artifacts of the changes made. Codebook is
useful for pairing a contributor with the original developer; however, the developer
may not have worked with the piece of code in years. A Hipikat program may provide
more information to the maintainer as it records the artifacts of why certain
design decisions were made when they were made using other tools like Bugzilla and
CVS. In both cases, our goal includes providing a summary of the topology of the
kernel repository through a visual tree.

Most visualizers provide a visual presentation of a certain aspect of a repository.
Fractal Figures \cite{Ambros2005} uses a unit square to represent a portion of a
project, then partitions the square based on the proportion of an author's
contributions to that portion of the project. EPOSee\cite{Burch2005} and Evolution
Radar\cite{Ambros2009} perform further analysis, determine which files are made
together, and what changes are made over a sequence of commits, though the goals
behind these projects is different.

Codebook, Hipikat, Fractal Figures, EPOSee, and Evolution Radar all work with data
from CVS repositories. Our goal is to provide information about Git repositories,
specifically the Linux master repository. Fewer tools are available for generating
visualizations and summaries of Git repositories, potentially due to the DAG model
used by Git.

Gource is a tool for providing an interactive timelapse of the state of a
repository\cite{Caudwell2010}. In the timelapse, it shows who contributes and what
type of contribution a developer is making. These contribution types are one of,
adding a file, removing a file, and changing a file. While the timelapse is
interesting to watch, it does not provide any additional explanation of the changes
actually being made, only the frequency that they are being made and who is making
them.

The current industry standard tool for Git repository visualization is Gitk. The
Gitk interface is built around the central DAG viewer. The DAG view displays the
commits on their respective branches, the author of the commit or merge, and the
date the commit or merge was authored. A user can select a single merge or commit
to view additional information. For both merges and commits, Gitk will present the
user with the Git log. If the user selected a commit, Gitk will also present the
user with the patch information and the names of the files edited. Gitk is unable to
provide the patches and filenames information in merges, as it is unable to
aggregate commit information in merges.

Our tool is primarily aimed at presenting the hierarchical structure of the Linux
git repository. We use tables for presenting the summarized information of the
commits and merges, but this information could also be presented in a graphical
form. Various graphical forms for displaying file and authorship data exist, the
principal forms being matrix views, city scapes, bar and pie charts, and networks
\cite{Eick2002}. Any of these data visualization metaphors are applicable to our
system.

\section{Conclusion}

Our tool shows promising results in working toward our goals, building a tool that
can more easily navigate the repository, and provide clearer explanations of the
changes made. The filtering by searches and the various tree visualizations enable
users to easily navigate through the kernel, finding the commits and merges that
pertain to what they are interested in. The tree visualizations are able to both
serve as a navigation piece, providing simpler navigation through the kernel, but
also demonstrate which commits contribute to which merges, providing further
explanation over the DAG. The tree model further improves the explanation by enabling
the tool to aggregate metadata at the merges instead of requiring a user to manually
aggregate the information. Our tool currently aggregates file information and the
modules edited, but can be extended to include authorship information and aggregated
patch information.

The tree model requires more information than what is provided by Git, requiring a
central server to monitor changes to the project. This breaks the design goals of
decentralized version control systems and makes the visualization tool specific to
the project.

While these are significant drawbacks for small projects, large projects will likely
have a more centralized work flow with a master repository where contributions will
eventually land, along with various servers that can watch this master repository
for changes. Given that the importance of clear visualizations of the repository
outweighs the cost of having a server monitor the state of the repository, the
tree-based model appears to provide clean visualizations and explanations of the
changes in the Linux kernel.

% \nocite{*}

\bibliographystyle{IEEEtran}
\bibliography{citations}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: