### 2017-06-11 paper/generalization.tex
Adding content for generalization and userstudy
This is just preliminary content for these sections, but it is heading
in the right direction.

The original algorithm uses a stack-like approach, which results in a depth-first
search of the tree. In the first pass, all of the commits that are in the
fist-parent position of the previous commit will be analyzed. Then following the
depth-first tactic, the tree that was earliest in the history of the repository will
be constructed, followed by the next earliest. This will be repeated until all of
the trees have been constructed and we reach the head of the master branch.

\evan{Provide a description of the new algorithm, and an DAG, and finally the
  resulting merge tree.}