### 2017-06-11 paper/userstudy.tex

We conducted a user study of \evan{NUMBER} participants to determine if there is a
statistically significant difference among users in performing certain tasks on
repositories between \textit{Linvis} and either gitk or the command-line git
interface. We were looking for improvements in accuracy, time performance, and
overall enjoyment among users when performing summarization and intra-tree
navigation, as these are the tasks that merge-trees are design to simplify.

\evan{still need to figure out who our intended user is... ``git user'' is a bit too
  broad.}

% TODO: Determine who the intended user is. "generic git user" is kind of broad.
% TODO: Add a note stating that we will refer to the primary branch as being the master branch

\subsubsection{Methodology}
\label{ssub:methodology}

\subsubsection{Commit Selection}
\label{ssub:commit_selection}

We selected two commits for the participant to draw and summarize by first
determining the size of tree in the first and second quartiles. We found that
between Linux 3.1 and 3.16, the size of the trees in the first quartile only
contained a single node, and that the size of the trees in the second quartile
resulted in trees up to seven nodes. The third quartile contained trees up to 51
nodes, and the fourth quartile contained one tree of 7217 nodes. It is worth
mentioning that the tree sizes in the fourth quartile increase very quickly, as the
next largest tree only contains 4708 nodes, and the third largest tree contains 2349
commits.

% TODO: Finish the paragraph on complexity
The complexity of the merge trees varies greatly.

The complexity of these trees varies greatly as well, . % TODO: Finish this sentence

We decided to work with one tree of one node, and one tree of seven nodes.

\subsubsection{Questions}
\label{ssub:questions}

We chose questions that were designed to test the two primary goals of merge trees,
compared with the results from gitk or command line git.

\subsubsection{Results}
\label{ssub:results}

\subsubsection{Discussion}
\label{ssub:discussion}