### 2017-06-12 paper/abstract.tex
Resizing text width
84-wide text doesn't fit split-screen on my little monitor.
Resizing it down to 72-wide.

Natively, git uses a directed acyclic graph model (DAG) with references
pointing from the most recent commit (TODO: most recent commit?, it's
not the \emph{HEAD} because that moves around) toward the initial commit
in the repository. This model is useful and necessary to the internal
functionality of the version control system, but does not provide users
with a clear nderstand of the events occuring in the repository. Due to
the direction of the link in the DAG, visualziation tools, like gitk,
are unable to provide summarized aggregation of the commits at a given
merge.  Furthermore, the visual DAG does not provide a clear explanation
of the events taking place in the repository.

Merge-Trees are a data structure derived from the DAG, that could
potentially alleiveate the issues with visualization built directly from
the DAG. The trees are defined such that the root is the merge into the
master branch, and the leaves are the individual commits. The references
point in both directed, from each node to the parent, and from the
parent to the children. Depending on the complexity of the merge, the
path from leaf to root may pass through multiple internal nodes.

In this paper, we verify that the Merge-tree model is able to simplify
the construction of a conceptual image of the parts of a repository, and
simiplify the summarization tasks through a user study. We will continue
by providing a more generalized approach to constructing merge-tress,
which will prune components of the DAG that are not necessary.

% \evan{Change this when we get the actual results, this is from the
% pilot study} We find that although the DAG view used by Gitk provides
% more information about the repository, in many cases it povides too
% much information and overwhelms the user.

% \evan{Move this} In some cases, where the structure of the merge tree
% is flat, that is to say that it has no inner merges, users are able to
% easily determine which commits are merged together, but are unable to
% determine which merge is the final merge is the root of the tree.