### 2017-06-15 paper/futurework.tex
Preliminary future work section

% Inter Navigation
% TODO: verify this with the study
While the merge-tree model provides users with a better conceptual image
of what is occuring within a single merge tree, we have not exprimented
or tested the potential gains of navigation between trees. At this time,
the Linvis search provides results grouped by merge-tree root instead of
as a single block of results. The results are generally similar, with
each merge tree being merged at a different point in time, or to a
different version of the kernel. A future extension of this may
incorporate some information about how a given cluster changes over
time. To do this, a technique would need to be developed to cluster
merge-trees into modules, and then ordering the merge-trees based on
time or release.

% TODO: ensure that linvis has been referenced/defined