### 2017-06-19 paper/abstract.tex
Minor rewording in abstract
Natively -> Internally; There are no other options

Fixes spelling of alleviate

Places TODO in a notice command.

Internally, git uses a directed acyclic graph model (DAG) with references
pointing from the most recent commit \notice{(TODO: most
    recent commit?, it's not the \emph{HEAD} because that moves
    around)} toward the initial commit in the repository. This model is
useful and necessary to the internal functionality of the version
control system, but does not provide users with a clear nderstand of the
events occuring in the repository. Due to the direction of the link in
the DAG, visualziation tools, like gitk, are unable to provide
summarized aggregation of the commits at a given merge.  Furthermore,
the visual DAG does not provide a clear explanation of the events taking
place in the repository.

Merge-Trees are a data structure derived from the DAG, that could
potentially alleviate the issues with visualization built directly from
the DAG. The trees are defined such that the root is the merge into the
master branch, and the leaves are the individual commits. The references
point in both directed, from each node to the parent, and from the
parent to the children. Depending on the complexity of the merge, the
path from leaf to root may pass through multiple internal nodes.

In this paper, we verify that the Merge-tree model is able to simplify
the construction of a conceptual image of the parts of a repository, and
simiplify the summarization tasks through a user study. We will continue
by providing a more generalized approach to constructing merge-tress,
which will prune components of the DAG that are not necessary.

% \evan{Change this when we get the actual results, this is from the
% pilot study} We find that although the DAG view used by Gitk provides
% more information about the repository, in many cases it povides too
% much information and overwhelms the user.

% \evan{Move this} In some cases, where the structure of the merge tree
% is flat, that is to say that it has no inner merges, users are able to
% easily determine which commits are merged together, but are unable to
% determine which merge is the final merge is the root of the tree.