### 2017-06-28 paper/results.tex
Adding results section

% vim:set et sw=2 ts=4 tw=72:
% Jun 24, 2017

In this section, we provide the raw aggregated results from the user
study and analysis on those results.

\subsection{Conceptual Questions Results}

% TODO: make plots showing distribution of times and responses

% Number of commits in the merge tree - Make a plot of this
% | text              | hash      | median |    mean | variance | Actual |
% |-------------------+-----------+--------+---------+----------+--------|
% | Number of commits | a3c1239eb |      4 | 19.1111 | 735.1111 |      1 |
% | Number of commits | cdbdd1676 |      4 |     7.8 | 136.8444 |      5 |
% | Number of merges  | a3c1239eb |      5 |  8.2727 |  53.6182 |      1 |
% | Number of merges  | cdbdd1676 |    3.5 |     5.4 |  50.2667 |      3 |
% Time taken to arrive at these results
% | text              | hash      | median(s) |  mean(s) | variance(s) |
% |-------------------+-----------+-----------+----------+-------------|
% | Number of commits | a3c1239eb |      10.0 |  49.9167 |   5952.0833 |
% | Number of merges  | a3c1239eb |       7.5 |  24.6667 |    884.4242 |
% | Number of commits | cdbdd1676 |      31.5 | 106.8333 |  54123.4242 |
% | Number of merges  | cdbdd1676 |      11.0 |     65.5 |  29798.8182 |

In Table~\ref{tab:conceptual_results}, we are able to see an overview of
the results from the conceptual questions.

\begin{table*}[htpb]
  \centering
  \caption{Results from the conceptual questions}
  \label{tab:conceptual_results}
  \begin{tabular}{ll|r|lrr|rrr}
    Question                      & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
    Number of commits in the tree & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
    Number of merges in the tree  & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
    Number of commits in the tree & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
    Number of merges in the tree  & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
  \end{tabular}
\end{table*}

Users were able to more closely estimate the number of commits and
merges in the larger tree, but generally took longer than the smaller
tree. The tree with a single node resulted in more variability in the
estimate of number of commits .

It should be noted that these questions were answered after spending
roughly ten minutes attesting to draw a picture that held the answers to
these questions.

\subsection{Summarization Question Results}