### 2017-07-05 paper/introduction.tex
Introduction: Linvis Introduction
Replacing introduction with the introduction from original paper.

% vim:set et sts=2 sw=2 ts=2 tw=72:

Git is a distributed version control system that is used in many
opensource and closedsource projects. Git enables many developers to
write code simultaneously, working on various components known as
branches. Git uses a directed acyclic graph model (DAG) for storing the
necessary information regarding the state of the code at any point in
time. Git refers to nodes of the DAG as comits, not distinguishing
between the code-carrying commits and the structural commits that bring
two branches together. In this paper, we will refer to the individual
commits as being commits, the structural merge commits as merges, and
refereing to all as repository events.

Between 50k and 70k commits are added to the Linux kernel per version
requiring maintainers of older versions of the kernel to sift through
thousands of commits and merges with tools that are unable to filter and
effectively visualize projects at the scale of the kernel. Older
versions of the kernel are used in embedded systems and mobile phones;
for security purposes, performance needs, and changing hardware
requirements, maintainers must be able to understand the changes being
made in the current version of the kernel in order to produce the
necessary patches for the older versions of the kernel. Current tools
and visualizations are based on the directed acyclic graph (DAG) of the
repository, showing all commits and merges in chronological order by
when the commit was authored, not by when it arrived in the official
Linux repository.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/gitk.png}
  \caption{
    The Gitk interface, centered at commit
    cdbdd1676a5379f1d5cbd4d476f5e349f445befe. The top left pane showing
    the DAG, the top center pane showing the authors, and top right pane
    showing the commit dates. The bottom-leftpane shows the full
    commit message and diff of the selected commit, and the bottom-right
    pane shows the list of the files modified by the selected commit.
  }
  \label{fig:gitk}
  %\vspace{-4mm}
\end{figure}

The DAG representation works in smaller projects; it enables users to
see when changes are made, when these changes are merged, how each
branch is interacting, and the point where a branch forks from the
master branch. In large modular projects, like the Linux kernel, the DAG
becomes a mess of merges and commits (Figure~\ref{fig:gitk}) losing its
visual meaning. In some cases, the repository is simply too large for
systems to generate a visualization; repositories stored on GitHub will
have a DAG visualzation to show the branches and forks of a
repositority, but is unable to display the visualization for projects at
the scale of Linux or even the Git Git
repository(Figure~\ref{fig:gitfail}).  Between 60k and 70k new commits
are created for the Linux project every year; according to previous
work\cite{German2015}, a commit takes a median of 30 days from the time
it is authored until it arrives in the official repository. The snapshot
of the kernel tomorrow may be different than the snapshot from today,
containing new commits authored in the past; distinguishing these new
commits from the commits in the snapshot from today is not trivial.

\begin{figure}
  \centering
  \includegraphics[width=0.8\linewidth]{figures/github_viewer.png}
  \caption{Github is unable to display the DAG of the Linux
    Project due to the size and complexity, which would normally
    show the relationship to other branches and forks.}
  \label{fig:gitfail}
  %\vspace{-2mm}
\end{figure}

This paper investigates the construction, implementation, user-based
evaluation, and generalization of a new model, the \mt model. We
begin with a description and algorithm for converting the DAG into a set
of merge trees. We continue by providing some details about the
web-based tool, called \tool\footnote{\tool is currently available at
  \url{http://li.turingmachine.org}}, implemented to test the
feasibility and potential capabilities of the \mt model. We
continue with a user study, to evaluate first, if the DAG is capable of
providing the users with a conceptual understanding of the repository
events that are related to a given commit, and second, determine if
there is a difference between the DAG and \mt model in performing
summarization tasks.

\subsection{Directed Acyclic Graph Model}
\label{sec:directed_acyclic_graph_model}

Git uses the DAG as the internal model for representing repositories.
This model functions perfectly, and is necessary, for the internal
functionality of Git. The DAG model allows for the reconstruction of
files and the working directory by recursively applying the changes, or
the diff, made in each repository event that is a parent of the current
event. We see an example DAG repository Figure~\ref{fig:repoDAG}.

Each repository event contains an ordered list of one or more parents,
except the initial commit, which has no parents. Commits have a single
parent, while merges will have at least two parents, the current branch,
and the branch being merged into the current branch. Git allows for the
merging of multiple branches simultaneously in octopus merges, where the
order of the parents represents the order that the branches were
specified for being merged.

The repository events in the DAG are immutable; once an event is
created, it can't be changed. Git allows operations to alter commits and
re-order them, but this will create a new commit with a new commit hash.
This property makes a repository event unable to record the traversal of
merges into the master branch of the repository. A commit cannot have
references in the direction of the merge that merges the commit, as this
reference will change, requiring the commit to be changed. While this
provides some challenge, git eliminates this issue of the DAG with the
\verb|git log --children| command, which traverses the DAG and inverts
the edges to provide the reference from parent to child. This
terminology will prove to be confusing through the rest of the paper,
when we are referring to the \mt form, the parent is the next node
toward the root. When referring to the DAG, the parent relationship is
in the opposite direction, from the root toward the branch-point.

Many current visualization systems base the visualization on the DAG,
because it is made readily available by git. The primary issue being
that the DAG was not designed with visualization and summarization tasks
in mind. Due to the direction of the links, the DAG model is unable to
provide a summarization of the contents of a merge; the DAG is unable to
show which branch is the master branch, and is unable to filter the
irrelevant information from the view.

\subsection{\mt Model}
\label{sub:merge_tree_model}

We use the \mt model to address these issues in the DAG\@. We
will describe the \mt model in detail in
Section~\ref{sec:mergetree}. The \mt reverses the direction of
the references of the DAG.\@ Starting with each merge into the master
branch, the algorithm recursively aggregates the statistics of each
commit. The tree is rooted at the merge into the master branch, and the
commits are the leaves.

\subsection{Implementation}
\label{sub:implementation}

We implement a tool, \tool, to test and verify the capabilities of the
\mt model. \tool provides search, summarization, and
visualization using the \mt model.

We provide users with search-based navigation. The search engine takes a
plain text query from the user and returns results based on similarity
to the query. The engine weighs various fields, including the commit
log, the author name, the commit hash, and the dates that the commit was
authored and committed. The results are aggregated by the \mt
with a link to the root at the top, and the repository events that match
the search query in a table below.  An example of a single \mt is shown
in Figure~\ref{fig:linvis_search_results}.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=1.0\linewidth]{figures/linvis/search_results.png}
  \caption{\tool search results for `net-next 2011', showing a single
    \mt group.}
  \label{fig:linvis_search_results}
\end{figure}

\tool uses seven tabs to show the information and visualizations for a
selected commit or merge. The message tab shows the full commit log
message. This does not include the diff. The files tab shows an
aggregated table of all files that were modified in a merge. It includes
metrics like the number of lines added, lines removed, total lines
modified, and the delta, summed across all commits in the merge tree
that modify this file. A small details drop-down button allows a user to
see exactly which commit makes the changes, as shown in
Figure~\ref{fig:linvis_files}. If the current repository event being
viewed is a commit, the aggregate views will only show modifications
made by the commit. The modules tab shows the modules modified in the
merge tree. Like the files tab, the modules tab uses a table to show the
name of the module, the number of commits that are in the merge tree
that work with the module, and a details button to provide the links to
those commits.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_files.png}
  \caption{Table showing the modified files in a merge, with the second
  entry expanded to show the commit that makes the changes.}
  \label{fig:linvis_files}
\end{figure}

The authorship tab is very similar to the files tab, but shows the
authorship information. It shows the sum of the number of lines added,
removed, modified, and the delta within the \mt. It also shows
the number of files that were modified by the author. The details tab is
organized slightly differently. Instead of organizing these results by
commit, the details are organized by file, showing which files were
modified by the author in this \mt (Figure~\ref{fig:linvis_authors}).

\begin{figure}[htpb]
  \centering
  \includegraphics[width=1.0\linewidth]{figures/linvis/linvis_authors.png}
  \caption{Table showing the authors who made changes in a merge. The
    entry for Randy Dunlap is expanded, showing the modifications to
    each file that were made by Randy.}
  \label{fig:linvis_authors}
\end{figure}

\tool implements three visualizations, a list tree, pack tree, and
Reingold-Tilford tree. The list tree is a series of indented textual
lists that can be easily searched with the built-in search in most
web-browsers. The list tree is rooted at the current event, if the event
is a commit, only that commit will be shown; conversely, if the selected
event is the root, the entire merge tree will be shown. The Pack-Tree or
Bubble tree, uses a set-based approach for showing the parent
relationship. The root node aggregates all of the events within the \mt
into the master branch. Thus, the root is the superset of all the other
nodes in the tree. The commits are shown as individual white circles,
containing no other events, as these are the leaves. This is shown in
Figure~\ref{fig:linvis_pack}. The selected event will be shown in bright
orange, commits in white, and merges in shades of blue. Darker shades of
blue indicate that the merge is deeper in the tree.

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/linvis/linvis_bubble.pdf}
  \caption{The pack tree visualization for a selected commit, shown in
    orange.}
  \label{fig:linvis_pack}
\end{figure}

The Reingold-Tilford tree is the classic tree visualization, with the
root at the top, leaves at the bottom, and the edges between them
showing the parent relationship. An example of the Reingold-Tilford tree
visualization is shown in Figure~\ref{fig:commit_1} and
Figure~\ref{fig:commit_2}.

\section{\mt Model}
\label{sec:mergetree}

The \mt model inverts the DAG, so that the reference point from
parent to child instead of child to parent. This is to enable
aggregation at merges. Given a merge, we are able to recursively
traverse all children of that merge, collecting the statistics at each
commit. The inverted DAG is then traversed starting at each merge into
the master branch. The merges into the master branch is the root of each
tree. The commits are the leaves of the tree. Starting at the merge into
the master branch, we recursively traverse each child of the current
node until we reach the branch point from the master branch, or a commit
that has a possible root merge that is closer in commit time to the
commit. As a consequence, commits will always take the shortest path, in
terms of time, to the master branch.

A short example: assume the commits represented in
Figure~\ref{fig:repoEvents} show the sequence of events in a repository.
In this case, commits are performed in various repositories and
branches. The DAG representation of the commits is shown in
Figure~\ref{fig:repoDAG}. Notice that the DAG loses information about
the master branch and the repository that the master branch is part of.
The \mt view of this DAG is visible in Figure~\ref{fig:repoTree}.
Note that the direction of the edges of the DAG have been inverted,
instead of pointing from the child to the ancestors, it points from the
ancestor to its successors, forming a path to the master branch. Also
note that the DAG has been simplified, showing only a single edge on the
path to master for any commit.

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
    \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7}
    \draw[shift={(\x + 0.5, -0.5)}, color=black] (0cm, 4cm) -- (0pt, -0.2cm);

    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
    \draw (10) edge[chartyellow, -stealth] (9);
    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
              edge[chartyellow, -stealth] (7);
    \draw (8) edge[chartmagenta, -stealth] (7);
    \draw (7) edge[chartyellow, -stealth] (5);
    \draw (6) edge[chartred, -stealth] (5);
    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
              edge[chartyellow, -stealth] (4);
    \draw (4) edge[-stealth] (1);
    \draw (3) edge[chartmagenta, -stealth] (2);
    \draw (2) edge[chartyellow, -stealth] (1);

    \node [draw=chartblue, below = 1.5cm of 1] (l1) {Master};
    \node [draw=chartyellow, right = 1.5cm of l1] (l2) {Repo A};
    \node [draw=chartred, right = 2.5cm of l2] (l3) {Branch of Repo A};
    \node [draw=chartmagenta, right= 2.5cm of l3] (l4) {Repo B};

        \foreach \x in {0, 1, 2, 3, 4, 5, 6, 7, 8}
    \node[shift={(\x, -0.6)}, color=black] {$t_\x$};
  \end{tikzpicture}
  \caption{Example of a sequence of events performed in different
    repositories. The horizontal axis represents time. Each horizontal
    section represents a different branch and/or repository. Each commit
    points to its ancestor.}
  \label{fig:repoEvents}
%\vspace{-3mm}
\end{figure}

% TODO: make everything one color
\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]
    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (12) edge[-stealth] (11) edge[chartyellow, -stealth] (10);
    \draw (11) edge[-stealth] (4) edge[chartyellow, -stealth] (9);
    \draw (10) edge[chartyellow, -stealth] (9);
    \draw (9) edge[chartmagenta, -stealth] (8) edge[chartred, -stealth] (6)
              edge[chartyellow, -stealth] (7);
    \draw (8) edge[chartmagenta, -stealth] (7);
    \draw (7) edge[chartyellow, -stealth] (5);
    \draw (6) edge[chartred, -stealth] (5);
    \draw (5) edge[chartmagenta, -stealth] (3) edge[chartyellow,-stealth] (2)
              edge[chartyellow, -stealth] (4);
    \draw (4) edge[-stealth] (1);
    \draw (3) edge[chartmagenta, -stealth] (2);
    \draw (2) edge[chartyellow, -stealth] (1);
  \end{tikzpicture}
  \caption{DAG representation of the commits represented in
    Figure~\ref{fig:repoEvents}. The DAG does not maintain the
    information about which repository a commit comes from. Furthermore,
    the DAG does not distinguish the master branch from other branches.}
  \label{fig:repoDAG}
\vspace{-3mm}
\end{figure}

\begin{figure}[htbp]
  \centering
  \begin{tikzpicture}[auto, on grid, semithick, state/.style={circle, text=black}]

    \draw[chartblue]
      (-0.5, -0.5) -- (8.5, -0.5) -- (8.5, 0.5) -- (-0.5, 0.5) -- (-0.5, -0.5);

    \node[state, draw=chartblue] (1) {1};
    \node[state, draw=chartyellow, above right= of 1] (2) {2};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 2] (3) {3};
    \node[state, draw=chartblue, right= 2cm of 1] (4) {4};
    \node[state, draw=chartyellow, above right=of 4] (5) {5};
    \node[state, draw=chartred, above right=of 5] (6) {6};
    \node[state, draw=chartyellow, right=of 5] (7) {7};
    \node[state, draw=chartmagenta, above right= 2cm and 1cm of 7](8) {8};
    \node[state, draw=chartyellow, right= 2cm of 7] (9) {9};
    \node[state, draw=chartyellow, right=of 9] (10) {10};
    \node[state, draw=chartblue, below right=of 9] (11) {11};
    \node[state, draw=chartblue, below right=of 10] (12) {12};

    \draw (2) edge[chartyellow, -stealth] (5)
          (3) edge[chartmagenta, -stealth](5)
          (5) edge[chartyellow, -stealth] (7)
          (7) edge[chartyellow, -stealth] (9)
          (6) edge[chartred, -stealth] (9)
          (8) edge[chartmagenta, -stealth] (9)
          (9) edge[chartyellow, -stealth] (11)
          (10) edge[chartyellow, -stealth] (12);

  \end{tikzpicture}
  \caption{\mt view of the commits  represented in
    Figure~\ref{fig:repoEvents} showing the path they followed to reach
    the master branch. In this model the successors of each commit
    represents the path followed by that commit to reach the master
    branch.}
  \label{fig:repoTree}
\vspace{-3mm}
\end{figure}

\subsection{Computing the \mt from the DAG of Linux}

Computing the \mt from a DAG for any repository may not be
possible; however, certain features of the development process of Linux
make it feasible to compute the \mt for the Linux repository.
First, the master branch of Linux is maintained by Linus Torvalds, and
only Linus has write access to it. We have verified this assertion in
previous research~\cite{German2015}. We have developed a heuristic that
is presented in Algorithm~\ref{fig:alg}. In short, the algorithm first
identifies the commits made directly to the master branch; whereafter it
recursively determines the shortest path (in terms of time), using the
DAG, from each commit to the master branch using the inverted DAG.

\begin{algorithm}
        \caption{Computing the \mt from Linux Git's DAG}\label{fig:alg}
        \begin{algorithmic}
                \Function{ComputeMergeTree}{DAG}: tree
                \State {\# Compute the tree from the DAG of Linux repository.}
                \State {\# Returns $Tree$, a graph containing every commit }
                \State {\# in DAG with the path it followed to master.}

                \State $head \gets \textit{Head of master of git repository}$
                \State $master \gets \textit{traverse DAG from head using }$
                \State \quad\quad\quad\quad $\textit{first ancestor until reaching root}$
                \State $nodes(Tree) \gets nodes(DAG)$
                \State \Function{distance2Master}{cid} : seconds
                \State {\# Helper function}
                \State {\# Recursively compute shortest distance to master}
                \State {\# setting cid's successor (next) in its way to master.}
                \State {\# This function should be memoized. Otherwise it}
                \State {\# would run in exponential time.}
                \If {\textit{cid in master}}
                \State \Return 0
                \EndIf
                \State    $d \gets 	\infty$
                \State {\# Traverse the inverted DAG}
                \For{$c \in children(cid, DAG)$}
                \If {$c \in master$}
                \State $d_1 \gets commitTime(c)-commitTime(cid)$
                \Else
                \State {$d_1 \gets distance2Master(c)$}
                \EndIf
                \If {$d_1 < d $}
                \State $next \gets c$
                \State  $d \gets d_1$
                \EndIf
                \EndFor
                \State {\# $c$ is the commit that follows $cid$}
                \State {\# in its way to master}
                \State add edge $(cid, next)$ to $Tree$
                \State \Return $d$
                \EndFunction

                \State {\# Compute the distance for each commit}
                \State {\# discarding result}
                \For{$c \in nodes(DAG)$}
                \State $distance2Master(c)$
                \EndFor
                \State \Return $Tree$
                \EndFunction
        \end{algorithmic}
\end{algorithm}

\subsection{Evaluation}

Merges that do not have merge conflicts provide information to verify
this heuristic. If a merge does not contain a conflict, it records a
summary of the commits that it merges. See Figure~\ref{fig:sampleMerge}
for an example. This summary contains a list of the first 20 non-merge
commits in the merge, including their one-line log description, the full
logs of the merge commits that merge this subset, and the total number
of non-merge commits in the merge.

\begin{figure}[htbp]
  \centering
  {\fontsize{7}{9}
    \begin{verbatim}
Merge: 8cbd84f fd8aa2c
Author: Linus Torvalds <torvalds@linux-foundation.org>
Date:   Tue Aug 10 15:38:19 2010 -0700

Merge branch 'for-linus' of git://neil.brown.name/md

* 'for-linus' of git://neil.brown.name/md: (24 commits)
md: clean up do_md_stop
[... edited for the sake of space]
md: split out md_rdev_init
md: be more careful setting MD_CHANGE_CLEAN
md/raid5: ensure we create a unique name for kmem_cache...
...
    \end{verbatim}}\vspace{-5mm}
  \caption{Example of how merges record a subset of commits being
    merged. The commit only shows the first 20 one-line summaries
    messages for the 24 non-merge commits it merged. The ending
    ``\ldots'' is part of the log and represents that other
    commits were merged.}
  \label{fig:sampleMerge}
\end{figure}

We used this information to evaluate the accuracy of the \mt
model extracted from the DAG\@. The method we followed started with the
extraction of the commit history up to July 20, 2016. We computed the
merge tree of every commit until then. Since Linus Torvalds mostly does
merging directly into master, we assumed that every merge by him is the
root of a \mt. As described above, the log of a merge-commit
usually contains the number of commits in the merge the first 20
summaries of commits being merged. We extracted merges by Linus Torvalds
using the command \mycode{log --merges --author='Torvalds} and compared
the number of commits according to the log with the number of commits in
the  \mt rooted in this commit. We also used the summaries of the
commits found in the merge (not necessarily all---see above) to make
sure those commits were in their corresponding merge tree. For example,
for the merge in Figure~\ref{fig:sampleMerge} we would expect that the
merge tree rooted at \mycode{8cbd84f} contains 24 commits, and the
one-line summaries corresponds to commits in that \mt. We also
inspected those with differences to make sure they were true errors.
The results can be summarized as follows:

\begin{itemize}

  \item

    Five merges were false-errors because their logs did not contain
    accurate information (were probably edited by hand). For example in
    \mycode{42a579a0f\ldots} one commit summary was missing (the line was
    empty), in \mycode {c55d267\ldots} the summaries were reordered.

  \item

    The heuristic correctly identified that 79 of Linus merges (between
    Jun 7, 2014 and Jun 2, 2014) were made to a branch (not master).
    This branch was merged at \mycode{3f17ea6d\ldots} which contained 6809
    commits.

  \item

    The heuristic worked perfectly until Sept 4, 2007, the earliest date
    that it could be verified.  Before this date, and until Dec 12,
    2006, merges did not include a summary of the commits they included,
    hence making it impossible to verify; during this period, however,
    we correctly identified the merges by Linus into master.

  \item

    Before Dec. 12, 2006 (1542 merges) our heuristic breaks due to the
    presence of a \textit{foxtrot} commit (\mycode{c436688\ldots}),
    which confounded the true master branch of a repository (see
    \url{http://bit-booster.blogspot.ca/2016/02/no-foxtrots-allowed.html}
    for a description of the issue).

\end{itemize}

In summary, of the merges after Sept. 4, 2007, our heuristic was correct
in 100\% of the 16,680 commits. It failed in 1,542 commits before Dec.
12, 2006 and in 836 it appears to be correct (Dec 7, 2006 to Sept 4,
2007).