### 2017-07-07 paper/lineval.tex

% vim:set et fenc=utf-8 ft=tex sw=2 ts=2 tw=72:

\newif\ifdraft
\drafttrue

\ifdraft
        \documentclass[draft]{IEEEtran}
        \def\baselinestretch{1}
        % \setlength{\marginparwidth}{1.5cm}
\else
        \documentclass[conference]{IEEEtran}
\fi

\input{preamble.tex}

\begin{document}

\title{\TheTitle}

\author{
        \IEEEauthorblockA{\TheAuthors}
        \IEEEauthorblockN{Department of Computer Science,
                University of Victoria, Canada.}
        \IEEEauthorblockA{Email: \TheEmails}
}
\maketitle

\begin{abstract}

  \input{abstract.tex}

\end{abstract}

\begin{IEEEkeywords}
  \TheKeywords
\end{IEEEkeywords}

\section{Introduction}

\input{introduction.tex}

\section{Merge-Tree model}
\label{sec:mergetree}

\input{model.tex}

\section{Visualizing the merge-tree of Linux}

The goal of \tool is to simplify the navigation of the kernel commit
information, specifically focusing on merges. This is done by leveraging
the merge-tree to inspect how commits are merged on the path to the
master repository.

\subsection{Use cases}

We designed \tool with two use-cases in mind, though a user may switch
between the cases as they work.

\noindent \textbf{Use-case 1: top-to-bottom
  approach}\label{sec:usecase1}\\ These are users that are maintaining a
section of the kernel and would like to pick a merge (including all the
commits that it merges) and merge it directly into their current
repository. This is useful for reducing the amount of re-implementation
work. For these users, it is important to have the ability to aggregate
metadata about files and modules being effected by the merge. Also, it
is important for these users to be able to navigate from the root of the
merge-tree toward the leaves.

\noindent \textbf{Use-case 2: bottom-to-top
  approach}\label{sec:usecase2}\\ These are users that start with a
known merge or commit and would like to see what other changes are being
made in commits that are in the same merge, including knowing the
merge-tree they belong to. This is useful to see what other commits are
related to the current commit and how they get collated into merges that
eventually end in the master branch. This is primarily for maintainers
that need to perform some specific cherry picking of commits. We must
provide these users a mechanism for navigating from a single commit
toward the master branch, allowing them to see other commits that might
be related to their original commit.

\subsection{Data Model}

In our visualizations, we leverage the merge-tree model described in
Section~\ref{sec:mergetree}. In this model each commit is either already
in the master branch, or is part of a tree which is rooted in the merge
that merged it into the master branch.  Each commit, whether a merge or
non-merge, has only one successor; the root of each tree has none as it
was made by Linus Torvalds directly into the master branch. Non-merge
commits contain the metadata for the changes made.  This metadata
includes the files changed, the lines added and removed from each file,
the author, the date the commit was merged into the merge that led to
being merged into the kernel, the date the commit was authored, the
patch, and the commit log. Merges contain less metadata, only storing
the author of the merge, the log, the commit date, and the author date,
and potentially, changes necessary to address conflicts during the
merge. The details of the model are outlined in \cite{German2015}.

\section{Design and Implementation}

\input{design.tex}

\section{Related Works}
\label{sec:related_works}

\input{relatedwork.tex}

\section{Controlled User Study}
\label{sec:study}

\input{study.tex}

\section{Discussion}
\label{sec:discussion}

\input{discussion.tex}

\section{Conclusion}
\label{sec:conclusion}

\input{conclusion.tex}

\balance

% \nocite{*}

\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: