### 2017-07-13 paper/study.tex
Correctness results for summarization tasks

The purpose of the study is to quantitatively and qualitatively evaluate
the effectiveness of \tool compared to the DAG-based visualizations
found in Gitk or the git command line. The study is broken into three
task-based sections; conceptual tasks, summarization tasks, and
participant opinions. The study was performed in June of 2017 in
Kingston ON, in a controlled environment running Ubuntu 14.04.
Participants were allowed to use both Gitk and the git command line
tools together, which we will refer to as the git tools or as Gitk. Then
they were asked to perform the same tasks using \tool.

\evan{Do I need to say where the study took place? I see it showing up
  in other papers, but it makes it kind of obvious about the lab I'm
  talking about, doesn't it?}

The primary goal of the study is to determine if \tool is more capable
of providing the participants of our study with a better conceptual
understanding of how a commit is integrated into the kernel, and a
better summarization of various metrics about the files and authors
involved with the integration. The conceptual questions are to determine
if the DAG visualization provides enough information to determine how a
commit, and any related commits, are merged into a project. \tool
displays this information directly in the form of a tree, so we will not
be performing a comparison on this section, only evaluating the results
from Gitk. The summarization tasks compare \tool and Gitk, with the
participants using one tool followed by the other on two commits from
merge-trees of differeing sizes. In both of these sections, the
participant is given a task that they are to find an answer to; we
evaluate the time it took, if the answer was correct, and how far from
the right answer they were. The user opinions provide us with a clear
comparison between the tools from the point-of-view of the participant,
providing us information about which aspects of each tool they prefered.

% TODO: Add the participant profile

\subsection{Tasks}
\label{sub:tasks}

% \dmg{use tasks, not questions. you are asking them to do something that requires an answer, that is different than a question}

\textbf{The conceptual tasks} are designed to gain better insight on the
comprehension of the DAG, and determine if the DAG is sufficient for
understanding what other commits are related to a given commit, and how
those commits are merged. The tasks are outlined in
Table~\ref{tab:conceptual_tasks}. The tasks are performed in order
using the git tools. T1 has the participant draw a diagram of the merge
tree, which will be used to answer the other two questions. This enables
us to visually see issues with comprehending the DAG visualizations.
Participants are given 10 minutes to perform this task, which also is
used to familiarize the participant with the git tools interface, if
they are not already, and with the set of commits they will be working
with. T2 and T3 are drawn directly from the diagram from T1, giving
quantitative metrics to measure the comprehension. T2 asks the
participant to verify the number of commits, and T3 asks for the number
of merges. The three tasks are performed on both commits before
continuing to the summarization tasks.

\begin{table*}[htpb]
  \centering
  \caption{Conceptual Tasks }
  \label{tab:conceptual_tasks}
  \begin{tabulary}{0.9\textwidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
    T2 & How many individual commits are related to this commit?\\
    T3 & How many merges are involved with merging this commit into the master branch?\\
    \bottomrule
  \end{tabulary}
\end{table*}

\textbf{The summarization tasks}, listed in Table~\ref{tab:summarization_tasks},
are designed to compare the ability of the participants to summarize
information about the two commits used in the study when using git tools
verses \tool. The tasks are broken into four sets based on what is being
summarized: merges, authorship, files, and modules. The order that the
participants perform each task is randomized within each task set, and
the order of the task sets are then randomized with the exception of the
merges task set which will always come first. We measure and evaluate
three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

\evan{Is this better, or do you still want me to move the information
  about the randomization to the bottom?}

\begin{table}[htpb]
  \centering
  \caption{Summarization Tasks}
  \label{tab:summarization_tasks}
  \begin{tabulary}{\linewidth}{LLL}
    \toprule
    Task Set   & Task & Description\\\midrule
    Merge      & T4   & What is the series of merges involved with merging this
    commit?\\
    & T5   & What other commits are merged?\\
    Authorship & T6   & How many authors are involved?\\
    & T7   & Who contributed the most changes?\\
    Files      & T8   & How many files were modified?\\
    & T9   & Which file had the most changes?\\
    Modules    & T10  & Which modules does this merge tree involve?\\
    \bottomrule
  \end{tabulary}
\end{table}

While the merge task set is part of the summarization task group, it
works less with the summarization of the commits and more with
comprehension of the models. The task is designed with two goals in
mind; first we want to measure and define which commits and which merges
the participant has defined to be in the merge tree; second to help
solidify which commits and merges will be used in the merge tree.

T4 provides a concrete series of merges that the participant believes to
be integrating the commit into the repository. T5 provides a concrete
set of commits that are related to the commit. We place these tasks
first as the answers to these will be necessary for the participant to
answer the tasks that follow, specifically finding the merge into the
master branch. T6 and T7 are related to the authorship at merges,
determining how many authors are involved and who made the most changes.
T8 and T9 are related to the files that were modified in the merge,
determining how many files were modified and which file had the most
changes. T10 is related to the modules that were modified in the merge.
Modules are not explicitly defined in git, but are a property of the
Linux repository that we noticed while looking at various commits. To
find the modules, take the text up to the first colon in the git log
preview. For the log preview \textit{``ALSA: kernel docs: fix
  sound/core/ kernel-doc''} the module is \textit{``ALSA''}, e.g.

The accuracy metrics are measured slightly differently between tasks due
to different requirements and structures in the answer. In task T4, both
the order and the merges are important. We define the distance to be the
edit distance with insertion, deletion, and transposition at unit cost.
The accuracy is the minimum edit distance necessary to make the answer
correct. T5, T9, and T10 are concerned with the set of elements in the
response; the order is not important. We use the Levenshtein distance,
with insertion, deletion, and substitution of unit cost, from the
provided answer to the correct answer. The answers to T6 and T8 are
single number values; the distance is measured as the absolute value of
the difference between the provided answer and the correct answer. T7
has only a single correct answer, thus the distance would either be zero
or one. For this reason, it is acceptable to omit the accuracy metric
from this task. The correctness metric is simply whether the answer was
correct or not; an accuracy of zero indicates a correct answer.

\textbf{The user-opinion} questions, listed in
Table~\ref{tab:opinion_questions}, are designed to determine which tool
provides a better user experience for summarization and comprehension
tasks, and what aspects of each assisted with their understanding. T11
asks for the preference of the participant, given that their goal is
comprehension and summarization. % TODO add T11 results
We recognize that neither tool is perfect, and participants may have
complaints about both tools, but are interested in what parts of each
tool assisted them with understanding the events in the repository. T12
is meant to address this. % TODO add T12 results

\begin{table}[htpb]
  \centering
  \caption{User-Opinion Questions}
  \label{tab:opinion_questions}
  \begin{tabulary}{0.9\linewidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T11 & Given these tasks again, which tool would you prefer?\\
    T12 & Which aspects of each tool did you like and why?\\
    \bottomrule
  \end{tabulary}
\end{table}

\subsection{Commit Selection}
\label{sub:commit_selection}

We include two commits in the study to detect differences between merge
tree sizes. The order that the commits are presented to each participant
is randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
nodes non-root nodes.
% TODO: get the number for this if there is time
75\% of the trees contain at least 51 nodes, and the largest tree
contains 7217 nodes.

From this information, we selected one tree with a single commit, and
one tree with seven other nodes to represent a small tree and a
medium-sized tree. A majority of the trees are flat, where all nodes
merge directly into the root node. Of the 8031 trees containing at least
seven nodes, only 593 contained at least a single internal merge node.
Trivially, trees with a single node cannot have any internal nodes, and
of the 624 trees with seven nodes, only 135 contained at least one inner
node. To increase the complexity of the medium-sized tree, we randomly
selected one of the 135 trees to represent the medium tree.

From the 2008 trees containing only a single node, we selected one at
random, and trivially selected the only commit in the tree. From the
trees containing seven nodes, we selected one randomly, then randomly
selected one of the commits in the tree. We selected commit
\emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree containing
a single node, which we will refer to as \comA, and commit
\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
seven nodes, which we will refer to as \comB. A visual representation
for these trees is shown in Figure~\ref{fig:study_commits}.

\begin{figure}[bpt]
  \centering
  \begin{tabular}{ m{1.5cm} m{3cm} }
    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
  \end{tabular}
  \caption{The merge trees used in the user study,
    containing one and seven commits, respectively.}
  \label{fig:study_commits}
\end{figure}

\subsection{Participant Profile}
\label{sub:participant_profile}

The study was conducted with 12 participants, all of whom were
masters, PhD, or post-doc researchers in the field of software
engineering. The participants had between 6 months and 10 years
experience with git, with the median being 3.5 years. Most of the
participants had additional experience with SVN and CVS.\@ One of the
participants in the study worked as a release engineer, studying merge
practices to determine the best way to merge branches while minimizing
the number of merge conflicts, on SVN repositories. The participants
worked with repositories ranging from around 10 commits up to 38000
commits, with the median being 350 commits. Two of the participants had
never collaborated with anyone in a repository, while the rest had some
experience with repositories being modified by multiple people, with the
most being 219. The median number of collaborators was four.

Five participants had only worked with personal and academic
repositories. Two participants had only worked with school and
professional repositories, and two had only worked with professional
repositories. One participant worked with personal, academic, and
professional repositories, one participant had only worked with personal
repositories, and one participant had worked with personal, academic,
and school repositories.

\evan{Do you want more information about the participants?}

All participants have had at least some experience with the idea of
version control, branching in repositories, and git. The participants
are all from the same lab, and each participant worked with both tools
in the study, thus, keeping the sample populations identical for both
tools, with some variation between participants in experience with
repositories.

\subsection{Study Procedure}
\label{sub:study_procedure}

The study begins with a short introduction to the \mt model, followed by
an introduction to Gitk and \tool. We provided two examples of
converting the DAG to the \mt. The first DAG was a flat merge, so the
corresponding merge tree did not have any internal nodes. The second was
not a flat merge, so the corresponding merge tree contained a single
internal node. Any questions about the model were answered at this time.
Following the introduction to the \mt model, we introduce Gitk and
\tool, providing information about what each pane shows in Gitk, and
what each tab does in \tool.

The conceptual tasks follow the introduction, followed by the
summarization, which are then followed by the user opinion. The study is
completed with three questions to enable us to know more about the
participants;

\begin{itemize}
  \item How long have you used git?
  \item For what kind of projects have you used git?
  \item How many commits, files, and collaborators were involved with
    the largest repository you have worked with?
\end{itemize}

To mitigate the order bias cased by using one tool before the other or
one commit before the other, we randomize both the tool and commit
between participants. The conceptual tasks are run twice, once with
\comA and once with \comB, in the same order for all participants using
the git tools.

\evan{This paragraph is kind of tricky to explain well. See if you can
  understand what I'm trying to say} The summarization tasks are run
four times, once with \comA and \tool, \comB and \tool, \comA and git
tools, and \comB and git tools. The participant will start with one
tool, and complete the tasks for both commits in that tool. The order of
the commits is the same as in the conceptual tasks. Once they complete
the summarization tasks for both \comA and \comB, they will switch to
the other tool and complete the tasks again on the commits in the same
order. The order of the tasks is consistent for all four runs.

The user opinions do not directly involve working with the tools or the
commits, so the ordering of the tools

% To mitigate the order bias caused by using one tool before the other,
% one commit before the other, and the various tasks, we randomize...

We use a script written in python 3.6.1 in order to assure that we do no
bias the randomization, and ensure that correct ordering is maintained
between tools, tasks, and commits. The script produces the entire script
for the study, so we only need to read it to correctly conduct the
study. Re-running the script generates the study for the next
participant. We use screen capture to record the audio and video. From
the screen capture, we measure the time, as well as recording the
responses in the study.

\evan{Should I include the python script? It would make explaining how
  the test was conducted 100\% clear.}

% TODO: maybe move the time measuring procedure from summ. to here

\subsection{Results}
\label{sec:results}

\evan{This looks weird not having text between the subsections... Should
  I put something here?...}

\subsubsection{Conceptual Tasks}
\label{sub:conceptual_tasks}

Table~\ref{tab:conceptual_results} shows a summary of the results from
conceptual tasks T2 and T3. The answer column contains the correct
answer to the corresponding task, while the median, mean, and variance
columns are with respect to the answers provided by the
participants.\evan{Is this better for describing the columns in the
  table?} The median(s), mean(s), and variance(s) columns are with
respect to the time taken, in seconds, by the participants to respond to
the tasks.

\evan{Should I add some of the drawings to the paper?}
The results from T1 did not vary greatly between the participants. The
drawings generally show something in the form of a list, starting from
the provided commit and traversing the master branch, indicating that
the participants were unable to determine which merge was in the master
branch. Many of the participants identified the next tag, v2.6.29-rc6,
to be ``the master branch'', however, a tag is not a branch. One
participant was able to draw the correct diagram for \comA directly
using the git command line, but rejected their response after inspecting
the DAG visualization provided by Gitk and proceeded to draw a list of
merges.

\begin{table*}[htpb]
  \centering
  \caption{Results from the conceptual questions}
  \label{tab:conceptual_results}
  \begin{tabular}{ll|r|lrr|rrr}
    Question & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
    T2       & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
    T3       & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
    T2       & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
    T3       & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
  \end{tabular}
\end{table*}

Users were able to more closely estimate the number of commits and
merges in the larger tree, but generally took longer than the smaller
tree. The tree with a single node resulted in more variability in the
estimate of number of commits.

It should be noted that these questions were answered after spending
roughly ten minutes attempting to draw a picture that held the answers
to these questions, so the times indicate how quickly the participant
was able to interpret their conceptual understanding.

\subsubsection{Summarization Tasks}
\label{sub:summarization_tasks}

% TODO: Add the null hypotheses for each question to the procedures
% TODO: Include the Wilcoxon Test and Cliff effect size test results

The results show ( Figure~\ref{fig:summarization_results} ) that in all
cases, except for T10, the number of instances of a correct result with
\tool is roughly the same as the number of incorrect instances with
gitk. \tool outperforms Gitk in correctness by a large margin. We use
the McNemar Chi-square test~\cite{McNemar1947} with continuity
correction to test the statistical significance of the hypothesis that
\tool improves the correctness of the participants. We use $\alpha =
0.05$, or a $95\%$ confidence level, therefore the $\chi^2$ test
statistic is 3.841.

\begin{figure*}[htpb]
  \centering
  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
  \caption{Results from the summarization tasks. The first row plots the
    correctness metric for each task. The second row plots the
    accuracy metric for each task. The third row plots the time metric
    for each task.}
  \label{fig:summarization_results}
\end{figure*}

\begin{table}[htpb]
  \centering
  \caption{Correctness Results}
  \label{tab:correctness_table}
  \begin{tabular}{crrc}
    \toprule
    Task & $\chi^2$ & $p$-value & Conclusion\\\midrule
    T4   & 13.07    & 0.0003    & Reject $H_0$\\
    T5   & 12.07    & 0.0005    & Reject $H_0$\\
    T6   & 14.06    & 0.0002    & Reject $H_0$\\
    T7   & 11.07    & 0.0009    & Reject $H_0$\\
    T8   & 13.07    & 0.0003    & Reject $H_0$\\
    T9   & 12.07    & 0.0005    & Reject $H_0$\\
    T10  & 4.00     & 0.0455    & Reject $H_0$\\
    \bottomrule
  \end{tabular}
\end{table}

The results for the correctness metric are summarized in
Table~\ref{tab:correctness_table}; the full results are as follows;

% TODO: add results for other metrics

\begin{itemize}
  \item

    T4; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 6                            & 0\\
                            & Incorrect & 15                           & 2\\
    \end{tabular}\\

  \item

    T5; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
    $12.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 5                            & 0\\
                            & Incorrect & 14                           & 4\\
    \end{tabular}\\

  \item

    T6; The $\chi^2$ statistic is $14.06$ with a p-value of $0.0002$.
    $14.06 > 3.841$ and $0.0002 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 5                            & 0\\
                            & Incorrect & 16                           & 2\\
    \end{tabular}

  \item

    T7; The $\chi^2$ statistic is $11.07$ with a p-value of $0.0009$.
    $11.07 > 3.841$ and $0.0009 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 8                            & 0\\
                            & Incorrect & 13                           & 2\\
    \end{tabular}\\

  \item

    T8; The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 6                            & 0\\
                            & Incorrect & 15                           & 2\\
    \end{tabular}\\

  \item

    T9; The $\chi^2$ statistic is $12.07$ with a p-value of $0.0005$.
    $12.07 > 3.841$ and $0.0005 < 0.05$, therefore we reject the null
    hypothesis.

    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}  & Correct   & 6                            & 0\\
      & Incorrect & 14                           & 3\\
    \end{tabular}

  \item

    T10; The $\chi^2$ statistic is $4$ with a p-value of $0.0455$.
    $4 > 3.841$ and $0.0455 < 0.05$, therefore we reject the null
    hypothesis. The p-value is very close to the $\alpha$ value in this
    case, so the conclusion is not as well defined.

    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
      & Incorrect & 8                            & 0\\
    \end{tabular}

\end{itemize}

\subsubsection{User Opinions}
\label{sub:user_opinions}

Among the 12 pariticipants, there was nearly unanimous agreement that
for conceptual understanding and summarization tasks, \tool was easier
to work with than Gitk. The participants cited the ability to abstract
information about the merge from the clean summarization tables and
simple visualizations. Three participants suggested that someone with a
professional understanding of Gitk and the git commandline may be to
extract the a good conceptual understanding from the DAG, and perform
the summarization tasks from the commandline. One of these three
participants said that they would prefer to have both tools available,
as they are able to complement each other.