### 2017-07-14 paper/study.tex
Editing first half of study

\section{Controlled User Study}
\label{sec:study}

We conducted a study to quantitatively and qualitatively evaluate the
effectiveness of \tool compared to the DAG-based visualizations found in
Gitk or the git command line. The study is broken into three task-based
sections; conceptual tasks, summarization tasks, and participant
opinions. The study was performed during June of 2017 in a controlled
environment running Ubuntu 14.04. To evaluate the DAG-based
visualizations, participants were allowed to use both Gitk and the git
command-line tools together, which we will refer to as Gitk. \tool was
used for evaluating \mt-based visualizations.

The primary goal of the study is two fold; first determining if the
DAG-based visualization is sufficient for conceptual understanding;
second comparing \tool and Gitk to determine which is more capable of
providing users with a summarization of various metrics involved with
integrating a commit into the repository. The conceptual questions are
to determine if the DAG-based visualizations (Gitk and git command line)
provide enough information to determine how a commit, and any related
commits, are merged into a project. \tool displays this information
directly in the form of a tree, so we will not be performing a
comparison on this section, only evaluating the results from Gitk. The
summarization tasks compare \tool and Gitk, with the participants using
one tool followed by the other on two commits from \mt{s} of
differing sizes. In both of these sections, the participant is given a
task that they are to find an answer to; we evaluate the time it took,
if the answer was correct, and how far from the right answer they were.
The user opinions provide us with a clear comparison between the tools
from the point-of-view of the participant, providing us information
about which aspects of each tool they preferred.

\subsection{Tasks}
\label{sub:tasks}

\begin{table*}[htpb]
  \centering
  \caption{Conceptual Tasks }
  \label{tab:conceptual_tasks}
  \begin{tabulary}{0.9\textwidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T1 & Draw a diagram showing how this commit was merged into the master branch, along with any other related commits\\
    T2 & How many individual commits are related to this commit?\\
    T3 & How many merges are involved with merging this commit into the master branch?\\
    \bottomrule
  \end{tabulary}
\end{table*}

\begin{table}[htpb]
  \centering
  \caption{Summarization Tasks}
  \label{tab:summarization_tasks}
  \begin{tabulary}{\linewidth}{LLL}
    \toprule
    Task Set   & Task & Description\\\midrule
    Merge      & T4   & What is the series of merges involved with merging this
    commit?\\
    & T5   & What other commits are merged?\\
    Authorship & T6   & How many authors are involved?\\
    & T7   & Who contributed the most changes?\\
    Files      & T8   & How many files were modified?\\
    & T9   & Which file had the most changes?\\
    Modules    & T10  & Which modules does this merge tree involve?\\
    \bottomrule
  \end{tabulary}
\end{table}

\textbf{The conceptual tasks} are designed to gain better insight on the
comprehension of the visualization in Gitk, and determine if it is
sufficient for understanding what other commits are related to a given
commit, and how those commits are merged. The tasks are outlined in
Table~\ref{tab:conceptual_tasks}. The tasks are performed in order using
Gitk. T1 has the participant draw a diagram of the merge tree,
which will be used to answer the other two questions. This enables us to
see issues with comprehending the DAG visualization. Participants are
given 10 minutes to perform this task, which also is used to familiarize
the participant with the Gitk interface, if they are not already, and
with the set of commits they will be working with. T2 and T3 are drawn
directly from the diagram from T1, giving quantitative metrics to
measure the comprehension. T2 asks the participant to verify the number
of commits, and T3 asks for the number of merges. The three tasks are
performed on both commits before continuing to the summarization tasks.

\textbf{The summarization tasks}, listed in
Table~\ref{tab:summarization_tasks}, are designed to compare the ability
of the participants to summarize information about the two commits used
in the study when using Gitk versus \tool. The tasks are broken into
four sets based on what is being summarized: merges, authorship, files,
and modules. The order that the participants perform each task is
randomized within each task set, and the order of the task sets are then
randomized with the exception of the merge task set which will always
come first. The order that the tools are used is kept consistent for a
participant, but randomized between participants. We measure and
evaluate three metrics for each task: correctness, accuracy, and timing.
Correctness measures whether the answer was correct or incorrect.
Accuracy measures how far from being correct the provided answer was.
Timing is the duration, in seconds, that the participant took to answer
the task. We begin timing the task at the end of the question, and stop
timing before the last modification to the answer. For this reason, it
is possible for times to overlap between tasks if the participant
changes their answer after answering another question. This will also
modify the answer used for measuring the accuracy. We measure the timing
from the screen capture.

While the merge task set is part of the summarization task group, it
focuses less with summarization of the commits and more with
comprehension of the model visualizations. The task is designed with two
goals in mind; first we want to measure and define which commits and
which merges the participant has defined to be in the merge tree; second
to help solidify which commits and merges will be used in the merge
tree.

T4 provides a concrete series of merges that the participant believes to
be integrating the commit into the repository. T5 provides a concrete
set of commits that are related to the commit. We place these tasks
first as the answers to these will be necessary for the participant to
answer the tasks that follow, specifically finding the merge into the
master branch. T6 and T7 are related to the authorship of a merge,
determining how many authors are involved and who made the most changes.
T8 and T9 are related to the files that were modified in the merge,
determining how many files were modified and which file had the most
changes. T10 is related to the modules that were modified in the merge.
Modules are not explicitly defined in git, but are a property of the
Linux repository that we noticed while looking at various commits. To
find the module name, take the text up to the first colon in the git log
preview. For the log preview \textit{``ALSA: kernel docs: fix
  sound/core/ kernel-doc''} the module is \textit{``ALSA''}, e.g.

The correctness metric is simply whether the answer was correct or not;
an accuracy of zero indicates a correct answer. The accuracy metrics are
measured slightly differently between tasks due to different
requirements and structures in the answer. In task T4, both the order
and the merges are important. We define the distance to be the edit
distance with insertion, deletion, and transposition at unit cost. The
accuracy is the minimum edit distance necessary to make the answer
correct. T5, T9, and T10 are concerned with the set of elements in the
response; the order is not important. We use the Levenshtein distance,
with insertion, deletion, and substitution operations of unit cost, from
the provided answer to the correct answer. The answers to T6 and T8 are
single number values; the distance is measured as the absolute value of
the difference between the provided answer and the correct answer. T7
has only a single correct answer, thus the distance would either be zero
or one. For this reason, it is acceptable to omit the accuracy metric
from this task.

Statistical significance testing is performed to verify that the results
are meaningful. We use $alpha = 0.05$ or a $95\%$ confidence level. The
McNemar $\chi^2$ test\cite{McNemar1947} with continuity correct is used
to test the statistical significance of the hypothesis that \tool
improves correctness of the participants. The $\chi^2$ threshold
statistic at $\alpha = 0.05$ is 3.841. The Wilcoxon\cite{Wilcoxon45}
test and Cliff's effect size\cite{Cliff93} are applied to the accuracy
and timing metrics to determine the significance of the difference
between \tool and Gitk. The Wilcoxon test is applied between commits to
determine if the size of the merge-tree effects timing and accuracy. The
null hypotheses and alternative hypotheses are stated in
Table~\ref{tab:null_alt_hypoths}.

\begin{table*}[h!]
  \centering
  \caption{Null and Alternative Hypotheses for Summarization Tasks\evan{Should I say exactly
      what the null hypotheses are referring to (counting authors,
      files, etc... or is this good enough? It's already a huge and
      redundant table) }}
  \label{tab:null_alt_hypoths}
  \begin{tabulary}{0.98\linewidth}{CCLL}
    \toprule
    Task & Metric             & Null Hypothesis                                                      & Alternative Hypothesis\\\midrule
    T4   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T4   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T4   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T4   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T5   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T5   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T5   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T5   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T6   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T6   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T6   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T6   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T7   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T7   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T7   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T7   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T8   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T8   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T8   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T8   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T9   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T9   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T9   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T9   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\

    T10   & \tiny{Commit}      & There is no difference on accuracy or timing between \comA and \comB & There is a difference on accuracy or timing between \comA and \comB \\
    T10   & \tiny{Correctness} & \tool does not effect the correctness of the participants            & \tool does effect the correctness of the participants\\
    T10   & \tiny{Accuracy}    & There is no difference between \tool and Gitk on accuracy            & There is a difference between \tool and Gitk on accuracy\\
    T10   & \tiny{Timing}      & There is no difference between \tool and Gitk on timing              & There is a difference between \tool and Gitk on timing\\
    \bottomrule
  \end{tabulary}
\end{table*}

\textbf{The user-opinion} questions, listed in
Table~\ref{tab:opinion_questions}, are designed to determine which tool
provides a better user experience for summarization and comprehension
tasks, and what aspects of each assisted with their understanding. T11
asks for the preference of the participant, given that their goal is
comprehension and summarization. % TODO add T11 results
We recognize that neither tool is perfect, and participants may have
complaints about both tools, but are interested in what parts of each
tool assisted them with understanding the events in the repository. T12
is meant to address this. % TODO add T12 results

\begin{table}[htpb]
  \centering
  \caption{User-Opinion Questions}
  \label{tab:opinion_questions}
  \begin{tabulary}{0.9\linewidth}{LL}
    \toprule
    Task & Description\\
    \midrule
    T11 & Given these tasks again, which tool would you prefer?\\
    T12 & Which aspects of each tool did you like and why?\\
    \bottomrule
  \end{tabulary}
\end{table}

\subsection{Commit Selection}
\label{sub:commit_selection}

We include two commits in the study to detect differences between \mt
sizes. The order that the commits are presented to each participant is
randomized. We analyzed the 15096 merge-trees from April 16th 2005 to
October 14th 2014, which corresponds to the releases Linux 2.6.12-rc3
and Linux 3.17-rc1. 25\% of the trees contain at most a single commit
not including the root, while 50\% of the trees contain at most seven
non-root nodes. 75\% of the trees contain at least 51 nodes, and the
largest tree contains 7217 nodes.

From this information, we selected one tree with a single commit, and
one tree with seven non-root nodes to represent a small tree and a
medium-sized tree. A majority of the trees are flat, where all nodes
merge directly into the root node. Of the 8031 trees containing at least
seven non-root nodes, only 593 contained at least a single internal
merge node. Trivially, trees with a single node cannot have any internal
nodes, and of the 624 trees with seven non-root nodes, only 135
contained at least one inner node. To increase the complexity of the
medium-sized tree, we randomly selected one of the 135 trees to
represent the medium tree.

From the 2008 trees containing only a single node, we selected one at
random, and trivially selected the only commit in the tree. From the
trees containing seven non-root nodes, we selected one tree randomly,
then randomly selected one of the commits in the tree. We selected
commit \emph{a3c1239eb59c0a907f8be5587d42e950f44543f8} from the tree
containing a single node, which we will refer to as \comA, and commit
\emph{cdbdd1676a5379f1d5cbd4d476f5e349f445befe} from the tree containing
seven nodes, which we will refer to as \comB. A visual representation
for these trees is shown in Figure~\ref{fig:study_commits}.

\begin{figure}[bpt]
  \centering
  \begin{tabular}{ m{1.5cm} m{3cm} }
    \includegraphics[height=0.5in]{figures/commits/1-commit.pdf} &
    \includegraphics[height=1in]{figures/commits/7-commits.pdf}\\
  \end{tabular}
  \caption{The merge trees used in the user study,
    containing one and seven non-root nodes, respectively.}
  \label{fig:study_commits}
\end{figure}

\subsection{Participant Profile}
\label{sub:participant_profile}

The study was conducted with 12 participants, all of whom were masters,
PhD, or post-doc researchers in the field of software engineering. The
participants had between 6 months and 10 years experience with git, with
the median being 3.5 years. Most participants had additional experience
with SVN and CVS.\@ One of the participants in the study worked as a
release engineer, studying merge practices to determine the best way to
merge branches while minimizing the number of merge conflicts, on SVN
repositories. The participants worked with repositories ranging from
around 10 commits up to 38000 commits, with the median being 350
commits. Two of the participants had never collaborated with anyone in a
repository, while the rest had some experience with repositories being
modified by multiple people, with the most being 219. The median number
of collaborators was four. Participants had most experience with
personal and academic repositories. Three of the twelve participants had
experience with professional repositories.

All participants have had at least some experience with version control,
branching in repositories, and git. The participants are from the same
lab, and each participant worked with both tools in the study, thus,
keeping the sample populations identical for both tools, with some
variation between participants in experience with repositories.

\subsection{Study Procedure}
\label{sub:study_procedure}

The study begins with a short introduction to the \mt model, followed by
an introduction to Gitk and \tool. We provided two examples of
converting the DAG to the \mt. The first example DAG was a flat merge,
so the corresponding \mt did not have any internal nodes. The
second was not a flat merge, so the corresponding \mt contained a
single internal node. Any questions about the model were answered at
this time. Following the introduction to the \mt model, we introduce
Gitk and \tool, providing information about what each pane shows in
Gitk, and what each tab does in \tool.

The conceptual tasks follow the introduction, followed by the
summarization, which are then followed by the user opinion. The study is
concluded with three questions to enable us to learn more about the
participants;

\begin{itemize}
  \item How long have you used git?
  \item For what kind of projects have you used git?
  \item How many commits, files, and collaborators were involved with
    the largest repository you have worked with?
\end{itemize}

To mitigate the order bias cased by using one tool before the other or
one commit before the other, we randomize both the order that the tools
are used, and the order that the commits are used, between participants.
The conceptual tasks are run twice, once with \comA and once with \comB,
in the same order for all participants using the Gitk.

\evan{This paragraph is kind of tricky to explain well. See if you can
  understand what I'm trying to say} The summarization tasks are run
four times, once with \comA and \tool, \comB and \tool, \comA and git
tools, and \comB and Gitk. The participant will start with one
tool, and complete the tasks for both commits in that tool. The order of
the commits is the same as in the conceptual tasks. Once they complete
the summarization tasks for both \comA and \comB, they will switch to
the other tool and complete the tasks again on the commits in the same
order. The order of the tasks is consistent over all four runs. The
questions in the user opinions section do not directly involve the tools
or commits.

We use a script written in python 3.6.1 in order to assure that we do
not bias the randomization, and ensure that correct ordering is
maintained between tools, tasks, and commits. The script produces the
entire script for the study, so we only need to read it to correctly
conduct the study. Re-running the script generates the study for the
next participant. We use screen capture to record the audio and video.
From the screen capture, we measure the time, as well as recording the
responses in the study.

\evan{Should I include the python script? It would make explaining how
  the test was conducted 100\% clear.}
\evan{I think it would be far too long then.}

\subsection{Results}
\label{sec:results}

\evan{This looks weird not having text between the subsections... Should
  I put something here?...}

\subsubsection{Conceptual Tasks}
\label{sub:conceptual_tasks}

Table~\ref{tab:conceptual_results} shows a summary of the results from
conceptual tasks T2 and T3. The answer column contains the correct
answer to the corresponding task, while the median, mean, and variance
columns are with respect to the answers provided by the
participants.\evan{Is this better for describing the columns in the
  table?} The median(s), mean(s), and variance(s) columns are with
respect to the time taken, in seconds, by the participants to respond to
the tasks.

\evan{Should I add some of the drawings to the paper?}

The results from T1 did not vary greatly between the participants. The
drawings generally show something in the form of a list, starting from
the provided commit and traversing the master branch, indicating that
the participants were unable to determine which merge was in the master
branch. Many of the participants identified the next tag, v2.6.29-rc6,
to be ``the master branch'', however, a tag is not a branch. One
participant was able to draw the correct diagram for \comA directly
from the git command line, but rejected their response after inspecting
the DAG visualization provided by Gitk and proceeded to draw a list of
merges.

\begin{table*}[htpb]
  \centering
  \caption{Results from the conceptual questions}
  \label{tab:conceptual_results}
  \begin{tabular}{ll|r|lrr|rrr}
    Question & Commit & Answer & Median & Mean  & Variance & Median(s) & Mean(s) & Variance(s)\\\hline\hline
    T2       & \comA  & 1      & 4      & 19.11 & 753.11   & 10.0      & 49.92   & 5952.08\\
    T3       & \comA  & 1      & 5      & 8.27  & 53.62    & 7.5       & 24.67   & 884.42\\\hline
    T2       & \comB  & 5      & 4      & 7.80  & 136.84   & 31.5      & 106.83  & 54123.42\\
    T3       & \comB  & 3      & 3.5    & 5.40  & 50.27    & 11.0      & 65.6    & 29798.82\\
  \end{tabular}
\end{table*}

Users were able to more closely estimate the number of commits and
merges in the larger tree, but generally took longer than the smaller
tree. The tree with a single node resulted in more variability in the
estimate of number of commits. It should be noted that these questions
were answered after spending roughly ten minutes attempting to draw a
picture that held the answers to these questions, so the times indicate
how quickly the participant was able to interpret their conceptual
understanding.

\subsubsection{Summarization Tasks}
\label{sub:summarization_tasks}

Table~\ref{tab:cross_commit_results} shows whether the merge-tree size
has an effect on the three studied metrics. An interesting observation
from the figure is that the correctness for T5 and T9 are effected by
the size of the merge-tree, but the accuracy is not. The correctness
results for T5 and T9 are shown in Figure~\ref{fig:correctness_t5_t9},
and the timing results for T7 are shown in Figure~\ref{fig:timing_t7}.

\begin{table}[htpb]
  \centering
  \caption{Cross-Commit Results\evan{Should I group the results by task
      or by metric? It's by metric right now}}
  \label{tab:cross_commit_results}
  \begin{tabular}{ccrl}
    \toprule
    Task & Metric      & $p$-value & Conclusion\\\midrule
    T4   & Correctness & 0.22      & Do not reject $H_0$\\
    T5   & Correctness & 0.04      & Reject $H_0$\\
    T6   & Correctness & 0.13      & Do not reject $H_0$\\
    T7   & Correctness & 0.06      & Do not reject $H_0$\\
    T8   & Correctness & 0.07      & Do not reject $H_0$\\
    T9   & Correctness & 0.04      & Reject $H_0$\\
    T10  & Correctness & 0.62      & Do not reject $H_0$\\

    T4  & Accuracy & 0.94 & Do not reject $H_0$\\
    T5  & Accuracy & 0.09 & Do not reject $H_0$\\
    T6  & Accuracy & 0.19 & Do not reject $H_0$\\
    T8  & Accuracy & 0.16 & Do not reject $H_0$\\
    T9  & Accuracy & 0.08 & Do not reject $H_0$\\
    T10 & Accuracy & 0.37 & Do not reject $H_0$\\

    T4  & Time & 0.77 & Do not reject $H_0$\\
    T5  & Time & 0.97 & Do not reject $H_0$\\
    T6  & Time & 0.90 & Do not reject $H_0$\\
    T7  & Time & 0.01 & Reject $H_0$\\
    T8  & Time & 0.87 & Do not reject $H_0$\\
    T9  & Time & 0.99 & Do not reject $H_0$\\
    T10 & Time & 0.68 & Do not reject $H_0$\\
    \bottomrule
  \end{tabular}
\end{table}

\begin{figure*}[htpb]
  \centering
  \includegraphics[width=\textwidth]{figures/userstudy/results.pdf}
  \caption{Aggregated results from the summarization tasks. The first row plots the
    correctness metric for each task. The second row plots the
    accuracy metric for each task. The third row plots the time metric
    for each task. The correctness results for T5, T9, and the timing
    results for T7 cannot be aggregated across commits, thus are
    omitted.}
  \label{fig:summarization_results}
\end{figure*}

\begin{figure*}[htpb]
  \centering

  \begin{tabular}{ m{6cm} m{6cm} }
    \includegraphics[width=6cm]{figures/userstudy/correctness/5.pdf} &
    \includegraphics[width=6cm]{figures/userstudy/correctness/9.pdf}
  \end{tabular}
  \caption{Unaggregated Correctness results for T5 and T9 \evan{Split
      this? Is it clear that between Commit 2 and Commit 1 is where the
    separation of the plots is?}}
  \label{fig:correctness_t5_t9}
\end{figure*}

\begin{figure}[htpb]
  \centering
  \includegraphics[width=0.8\linewidth]{figures/userstudy/time/7.pdf}
  \caption{Unaggregated Timing results for T7}
  \label{fig:timing_t7}
\end{figure}

\begin{table*}[htpb]
  \centering
  \caption{Summarized Correctness, Accuracy, and Timing results for the summarization tasks}
  \label{tab:summarization_table}
  \begin{tabular}{ccc}
    % Correctness
    \begin{tabular}{crrc}
      \toprule
      Task                  & $\chi^2$ & $p$-value  & Conclusion\\\midrule
      T4                    & 13.07    & 0.0003     & Reject $H_0$\\
      \multirow{2}{*}{T5}   & 5.14     & 0.0233     & Reject $H_0$\\
                            & 5.14     & 0.0233     & Reject $H_0$\\
      T6                    & 14.06    & 0.0002     & Reject $H_0$\\
      T7                    & 11.07    & 0.0009     & Reject $H_0$\\
      T8                    & 13.07    & 0.0003     & Reject $H_0$\\
      \multirow{2}{*}{T9}   & 4.17     & 0.0412     & Reject $H_0$\\
                            & 6.13     & 0.0133     & Reject $H_0$\\
      T10                   & 4.00     & 0.0455     & Reject $H_0$\\
      \bottomrule
    \end{tabular}
    &

    % Accuracy
    \begin{tabular}{clrc}
      \toprule
      Task & $p$-value            & Delta Est. & Conclusion\\\midrule
      T4   & $7.8 \times 10^{-5}$ & -0.61      & Reject $H_0$\\
      T5   & $7.2 \times 10^{-5}$ & -0.63      & Reject $H_0$\\
      T6   & $1.2 \times 10^{-5}$ & -0.69      & Reject $H_0$\\
      T8   & $5.8 \times 10^{-4}$ & -0.54      & Reject $H_0$\\
      T9   & $0.009$              & -0.42      & Reject $H_0$\\
      T10  & $0.002$              & -0.35      & Reject $H_0$\\
      \bottomrule
    \end{tabular}
    &
    % Timing
    \begin{tabular}{clrc}
      \toprule
      Task                 & $p$-value & Delta Est. & Conclusion\\\midrule
      T4                   & 0.0007    & -0.58      & Reject $H_0$\\
      T5                   & 0.0142    & -0.42      & Reject $H_0$\\
      T6                   & 0.0168    & -0.41      & Reject $H_0$\\
      \multirow{2}{*}{T7}  & 0.2586    & -0.29      & Do not reject $H_0$\\
                           & 0.0780    & -0.43      & Do not reject $H_0$\\
      T8                   & 0.0009    & -0.57      & Reject $H_0$\\
      T9                   & 0.0002    & -0.63      & Reject $H_0$\\
      T10                  & 0.0001    & -0.66      & Reject $H_0$\\
      \bottomrule
    \end{tabular}
  \end{tabular}
\end{table*}

Summaries for correctness, accuracy, and timing are shown in
tables~\ref{tab:summarization_table}.
The full results are as follows;

\begin{itemize}
  \item \emph{T4}: What is the series of merges involved with merging this
    commit?

    \textbf{Correctness:}
    The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
    hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
                            &           & \multicolumn{2}{c}{Linvis}\\
                            &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 6                            & 0\\
                            & Incorrect & 15                           & 2\\
    \end{tabular}
    \vspace{3mm}

    \tool improves the correctness when determine the series of merges
    involved with merging a commit.

    \textbf{Accuracy:} The p-value is $7.8 \times 10^{-5}$, which is
    less than 0.05; therefore we reject the null hypothesis. The Delta
    estimate is -0.61, which indicates a large effect size. \tool
    improves accuracy when determining the series of merges involved
    with merging a commit.

    \textbf{Timing:}
    The p-value is $0.0007$, which is less than 0.05; therefore we
    reject the null hypothesis. The delta estimate is -0.58, which
    indicates a large effect size. \tool decreases the time taken to
    determine the series of merges involved with merging a commit.

    Overall, \tool is able to help users determine the series of merges
    that a commit takes more quickly and more accurately than Gitk.

  \item \emph{T5}: What other commits are merged with this commit?

    \textbf{Correctness:}
    The correctness results are split for this task. For \comA, the
    $\chi^2$ statistic is $5.14$ with a $p-value$ of $0.02334$. $5.14 >
    3.841$ and $0.02334 < 0.05$, therefore we reject the null
    hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
                             &           & \multicolumn{2}{c}{Linvis}\\
                             &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}  & Correct   & 4                            & 0\\
                             & Incorrect & 7                            & 0\\
    \end{tabular}
    \vspace{3mm}

    The $\chi^2$ statistic for \comB is $5.14$ with a p-value of
    $0.02334$. $4.14 > 3.841$ and $0.02334 < 0.05$, therefore we reject
    the null hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
                             &           & \multicolumn{2}{c}{Linvis}\\
                             &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}  & Correct   & 1                            & 0\\
                             & Incorrect & 7                            & 4\\
    \end{tabular}
    \vspace{3mm}

    While there is a difference in correctness between different-sized
    merge-trees, \tool improves correctness of determining the other
    commits related to a commit in trees of varying sizes. The
    difference in the two commits stem from the participants that did
    not change outcomes between the tools.

    \textbf{Accuracy:}
    The p-value is $7.2 \times 10^{-5}$, which is less than 0.05;
    therefore we reject the null hypothesis. The Delta estimate is
    -0.63, which indicates a large effect size. \tool improves
    accuracy when determining other commits related to a commit.

    \textbf{Timing:}
    The p-value is $0.0142$, which is less than 0.05; therefore we
    reject the null hypothesis. The delta estimate is -0.58, which
    indicates a large effect size. \tool decreases the time taken to
    determine other commits related to a commit.

    Overall, \tool helps users determine other commits that are merged
    with another commit more quickly and more accurately.

  \item \emph{T6}: How many authors are involved in this merge?

    \textbf{Correctness:}
    The $\chi^2$ statistic is $14.06$ with a p-value of $0.0002$.
    $14.06 > 3.841$ and $0.0002 < 0.05$, therefore we reject the null
    hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 5                            & 0\\
      & Incorrect & 16                           & 2\\
    \end{tabular}
    \vspace{3mm}

    \tool improves correctness when determining how many authors are
    involved with a merge.

    \textbf{Accuracy:} The p-value is $1.2 \times 10^{-5}$, which is
    less than 0.05; we reject the null hypothesis. The delta estimate is
    -0.69, which indicates a large effect size. \tool improves accuracy
    when determining how many authors are involved with a merge.

    \textbf{Timing:} The p-value is 0.0168, which is less that 0.05; we
    reject the null hypothesis. The delta estimate is -0.41, which
    indicates a medium effect size. \tool decreases the time taken to
    determine the number of authors involved in a commit.

    Overall, \tool assists users with determining the number of authors
    involved in a merge more quickly and more accurately.

  \item \emph{T7}: Who contributed the most changes to this merge?

    \textbf{Correctness:}
    The $\chi^2$ statistic is $11.07$ with a p-value of $0.0009$.
    $11.07 > 3.841$ and $0.0009 < 0.05$, therefore we reject the null
    hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 8                            & 0\\
      & Incorrect & 13                           & 2\\
    \end{tabular}\\
    \vspace{3mm}

    \tool improves the correctness of participants when determining the
    author that contributed the most changes to a merge.

    \textbf{Timing:} The null hypothesis is not rejected in the case of
    \comA, but is rejected in the case of \comB. This indicates that
    \tool is able to help users determine who made the most
    contributions to a merge in a large tree but not in a small tree.

    Overall, \tool improves the correctness for trees of various sizes,
    but only has an effect on the length of time taken when the tree is
    large.

  \item \emph{T8}: How many files were modified in this merge?

    \textbf{Correctness:}
    The $\chi^2$ statistic is $13.07$ with a p-value of $0.0003$.
    $13.07 > 3.841$ and $0.0003 < 0.05$, therefore we reject the null
    hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk} & Correct   & 6                            & 0\\
      & Incorrect & 15                           & 2\\
    \end{tabular}\\
    \vspace{3mm}

    \tool improves the correctness of users when determining the number
    of files that were modified in a merge.

    \textbf{Accuracy:} The p-value is $5.8 \times 10^{-4}$, which is
    less than 0.05; we reject the null hypothesis. The delta estimate is
    -0.54, indicating a large effect size. \tool improves accuracy when
    determining how many files were modified at a merge.

    \textbf{Timing:} The p-value is 0.0009, which is less than 0.05; we
    reject the null hypothesis. The delta estimate is -0.57, indicating
    a large effect size. \tool decreases the time taken to determine the
    number files modified in a merge.

    Overall, \tool assists users to determine the number of files
    modified in a merge more quickly and accurately.

  \item \emph{T9}: Which file(s) had the most changes in this merge?

    \textbf{Correctness:}
    The correctness results are split for this task. For \comA, the
    $\chi^2$ statistic is $4.17$ with a p-value of $0.0412$. $4.17 >
    3.841$ and $0.0412 < 0.05$, therefore we reject the null hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
                             &           & \multicolumn{2}{c}{Linvis}\\
                             &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}  & Correct   & 5                            & 0\\
                             & Incorrect & 6                            & 0\\
    \end{tabular}
    \vspace{3mm}

    The $\chi^2$ statistic for \comB is $6.125$ with a p-value of
    $0.0133$. $6.125 > 3.841$ and $0.0133 < 0.05$, therefore we reject
    the null hypothesis.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
                             &           & \multicolumn{2}{c}{Linvis}\\
                             &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}  & Correct   & 1                            & 0\\
                             & Incorrect & 8                            & 3\\
    \end{tabular}
    \vspace{3mm}

    While there is a different in correctness between different-sized
    \mt{s}, \tool improves the correctness when determining which files
    had the most changes in trees of varying sizes. The difference in
    the two commits stem from the participants that did not change
    outcomes between tools.

    \textbf{Accuracy:} The p-value is $0.009$, which is less than
    $0.05$; we reject the null hypothesis. The delta estimate is -0.42,
    indicating a medium effect size. \tool improves accuracy when
    determining which file had the most changes.

    \textbf{Timing:} We reject the null hypothesis; the delta estimate
    is -0.63, indicating a large effect size. \tool decreases the time
    taken to determine which file had the most changes.

    Overall, \tool is able to assist users determine which file had the
    most changes more quickly and accurately.

  \item \emph{T10}: Which modules does this \mt involve?

    \textbf{Correctness:}
    The $\chi^2$ statistic is $4$ with a p-value of $0.0455$.
    $4 > 3.841$ and $0.0455 < 0.05$, therefore we reject the null
    hypothesis. The p-value is very close to the $\alpha$ value in this
    case, so the conclusion is not as well defined.

    \vspace{2mm}
    \begin{tabular}{cc|rr}
      &           & \multicolumn{2}{c}{Linvis}\\
      &           & Correct                      & Incorrect\\\hline
      \multirow{2}{*}{Gitk}   & Correct   & 14                           & 1\\
      & Incorrect & 8                            & 0\\
    \end{tabular}
    \vspace{3mm}

    \tool improves correctness when determining the modules involved at
    a merge.

    \textbf{Accuracy:} The null hypothesis is rejected; the delta
    estimate is -0.35, indicating a medium effect size. \tool improves
    accuracy when determining which modules are involved with a merge.

    \textbf{Timing:} The null hypothesis is rejected; the delta estimate
    is -0.66, indicating a large effect size. \tool decreases the time
    taken to determine the modules involved with a merge.

    Overall, \tool assists users determine which modules are involved
    with a merge more quickly and accurately.

\end{itemize}

\tool is able to assist users with summarizing various aspects of a
merge more quickly and accurately than Gitk. In two tasks, T5 and T9,
correctness was affected by the size of a tree, but in both trees, \tool
was able to provide a statistically significant improvement to the
correctness. Only in one task, T7, was timing affected by the size of a
tree; \tool did not significantly improve the time taken to answer T7 on
the smaller tree, but did improve the time taken on the larger tree.

\subsubsection{User Opinions}
\label{sub:user_opinions}

Among the 12 participants, there was nearly unanimous agreement that for
conceptual understanding and summarization tasks; \tool was easier to
work with than Gitk. The participants cited the ability to abstract
information about the merge from the clean summarization tables and
simple visualizations. Three participants suggested that someone with a
professional understanding of Gitk and the git command-line may be to
extract a conceptual understanding from the DAG visualization and
perform the summarization tasks. One of these three participants said
that they would prefer to have both tools available, as they are able to
complement each other.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "lineval.tex"
%%% End: