### 2017-07-19 response/response.txt
adding response letter

We would like to thank the reviewers for their extensive comments and
analysis. The paper has been considerably improved as a results of the
review.

The paper has been heavily modified, shifting focus from the
generalization of the algorithm toward the comparison user study between
Linvis and Gitk.

Linvis has since had minor updates. We removed most of the search
interface, replacing it with a full-text search engine that searches by
author name, filename, commit hash, commit and authorship date, and
keywords from the commit logs. The results of the search engine are
aggregated together by merge-tree.

The responses to Reviewer 1:

      While the technical descriptions of the algorithms are detailed and
      comprehensive, they need to be re-written for better clarification. The
      reading of Section III.B is specially hard. It is a complicated
      algorithm that demands a lot of attention from the reader. As such, in
      order to keep the reader interested, the authors should improve the
      description by adding a diagram or a flowchart. A simple example that
      the reader can follow along with the authors would also be a great
      improvement.

We provided a further description of the original algorithm followed by
a step-by-step example starting at commit 5 from the example in Figure 3
to help clarify things in Section III.A.

The generalized algorithm that was presented in section III.B was moved
to a separate section (Section IX). We re-worded and extended the
description and provided an example starting with the merge commit 11
(the algorithms run in opposing directions on the DAG) from the example
shown in Figure 3.

      The testing and evaluation of the generalized version of the algorithm
      must also be improved. Right now the test suite and the results are
      not concretely described and must be better clarified.

This is not applicable anymore. The generalization has become a
discussion point in Section IX. The proposed algorithm is a possible
means of improving the runtime performance by pruning the DAG to contain
a smaller superset of the repository events necessary for computing the
merge-tree of a given merge. The test suite was not a comprehensive test
of all possible merges, but tested some of the issues that we
encountered while designing it and the original algorithm.

    While the tree-based visualizations are well described, they are not
    novel visualizations and simply use abstractions that are already
    available from previous work. As such, the practical use of these
    visualizations in the context of the paper should be better described.
    The authors are advised to add at least one fully described use case,
    as a step-by-step that can be followed by the reader, showing how the
    visualizations are actually used to obtain useful insight from the
    merge-trees.

The intent is less focused on providing a new metaphor for trees, and
more focused on providing a new visualization of git repositories. The
visualization is focused on showing how a commit is integrated into the
master branch.

We conducted a user study to verify that the goals of the model and tool
are met with the visual metaphors and tables provided. The study is
broken down into what we see as being the two major issues with the DAG;
conceptual understanding and summarization.

See section VII.

Responses to the Attached PDF:

    Introduction
    The last paragraph of the intro is a bit confusing. It is not clear
    whether the authors mean that the algorithm was extended to work with
    more than one repository, or with repositories other than the Linux
    kernel (or both at the same time, which is what I think is the right
    answer). Also, they talk about an extension, explain its goals, but
    not what that extension is, i.e., how the original algorithm is
    actually extended to reach the goals (that should be done here only
    briefly, of course). A clearer explanation here would help motivate
    the reader to continue reading.

The generalization has been moved to the discussion portion of the paper
to signify that it is less of a contribution, and more of an idea. The
extension section is introduced with

"
The visualizations of the Merge-Tree model assist users with
understanding how commits are integrated. We continue by discussing
possible methods for improving the algorithm to directly produce the
combined DAG Merge-Tree without a post-processing step and pruning the
DAG to improve the performance and make it feasible to compute over the
internet. Furthermore, we look to generalize the algorithm to work with
arbitrary repositories.
"

To address the confusion surrounding the type of generalization, we
specify it in the last sentence of the first paragraph of section IX.

"Furthermore, we look to generalize the algorithm to work with arbitrary
repositories."

    In page 2, column 2, line 28, the authors introduce the concept of an
    “event”. By reading further, one realizes that “events” are either
    commits or merges. However, in previous paragraphs, the authors
    regarded this as “merge commits” and ”non-merge commits”. A
    clarification of these two terminologies would be useful to make the
    text clearer.

The terminology surrounding events, commits, and merges has been
clarified in the last sentence of the first paragraph of the Background
Section.

"In this paper, we will refer to the individual code-carrying commits as
commits, the structural merge commits as merges, and refer to both types
without discrimination as repository events."

    In Figure 3, legend: The sentence “Each horizontal section” is not
    totally clear. One way of improving the figure would be to add a
    separator between branches/repos, effectively turning the image into a
    matrix. Or simply explaining more clearly that nodes are arranged in
    rows. Also, the authors should mention that the colors also represent
    branches/repos.

The caption has been clarified and we state that the branches and
repositories are color-coded.

"
An example sequence of events performed in different repositories. The
horizontal axis represents time. The branches and repositories are
aligned horizontally, and color-coded. Each commit points to its parent.
The initial commit is at time t0, and the head is at t8.
"

    In Figure 4: the authors mention that the DAG loses information about
    which repository or branch each commit is performed in, but the figure
    retains the nodes’ colors, which contradicts that statement. While the
    reader comprehends the concept, removing the colors of the figure
    would probably communicate the message better.

The color was removed from figure four to clarify that branch and
repository information is removed in the DAG.

    Figures 3, 4 and 5 represent a sequence of steps in the explanation of
    the algorithm, so it would be significantly better for the reader if
    they were all in the same page.

The figures have been placed on the same page, and are in the same
column.

    Constructing the Merge Tree
    A small description of the children function on Algorithm 1 would be
    useful, as it searches for children in the opposite way than a normal
    depth-search algorithm does. A normal depth-search would traverse the
    nodes to which cid points to (its ancestors in git), but children does
    the opposite: it traverses the nodes which point to cid

The children function is the list of (DAG) children for the commit. I am
unsure of how to make this more explicit.

    There is apparently a problem when calculating d in Algorithm 1. It is
    only updated when a commit is directly connected to a master commit,
    i.e., it has a child c in master. As an example, consider first node 9
    from Fig. 4. It has a child in master (node 11), so its distance is
    d(9) = commitTime(11) – commitTime(9). Now let’s consider node 8. Its
    only child is node 9, which is not in master. Thus, it recursively
    calls distance2Master(9), which returns d(9). Since d(9) is smaller
    than infinity, then d(8) = d(9). But shouldn’t it be d(8) =
    commitTime(8) – d(9)? The problem becomes worse when we consider node
    7. It has two children, nodes 8 and 9. If d(8) = d(9), then how can
    node 7 know which of its children to choose? By looking at Figure 5 it
    obviously chose node 9, which seems to be the right choice, but was it
    an arbitrary choice?

Thank you for identifying this issue. We missed a few details of the
algorithm as implemented when describing it in the paper. We have
carefully followed our implementation to re-describe the algorithm more
accurately. Furthermore, a full example stepping through the algorithm
has been provided, starting at commit 5. We have worked to fully clarify
the algorithm and how it works. We are also making the source code
available in the replication package.

    Section III.B starts abruptly without a clear motivation on what the
    Generalized Merge Tree Construction is supposed to do. The authors
    should better motivate the jump from the Linux kernel tree to the
    generalized one. Also, it seems that the authors have used “branch”
    and “commit” interchangeably, without a clear distinction. The
    sentence “There is still some ambiguity in which child is the correct
    child to link to for a branch after phase 1”, for example, is not
    clear. In the sentence “The order of children at a branch node is
    based on the length of the path to that node, not the order that the
    nodes show up in as parents.”, what is a “branch node”, and what does
    the order of its children impact? Such concepts were not introduced
    before this section, and as such it makes the reading very hard. I
    suggest the authors start this section with a more general and
    “lightweight” introduction of concepts and ideas before going fully
    into technical details.

The section has been removed and replaced with a new section for
discussing the generalization, Section IX.

The first sentence motivates generalizing the algorithm as we have
quantifiable evidence that the merge trees help users. Therefore a
generalization is the next step.

"
The visualizations of the Merge-Tree model assist users with
understanding how commits are integrated into a project. We
continue by discussing possible methods for improving the
algorithm to produce the combined DAG Merge-Tree without
a post-processing step and pruning the DAG to improve the
performance and make it feasible to compute over the internet.
Furthermore, we look to generalize the algorithm to work with
arbitrary repositories.
"

    In Algorithm 2, what is a parent? Are we dealing with the normal DAG
    or the inverted DAG? Apparently neither, because Algorithm 2 does not
    mention any symbol called DAG. But if Algorithm 2 starts from the root
    commit, and according to Section II the root has no ancestors, who are
    the parents being accessed in line 6 of the algorithm (and everywhere
    else)? It would be logical to assume that Algorithm 2, starting from
    the root, would have no parents and thus only run once and exit. I
    assume that by “root” the authors mean any master commit (so it
    would be the root of one merge-tree, and not of the whole
    repository).

This is addressed by specifying the terminology surrounding parents and
children in the first paragraph of the Merge-Tree model section.

"In this model, not only has the DAG been inverted (and simplified), but
the entire notion of parent-child has been reversed. Due to this
property, the terminology will be different depending on the model we
are referring to; when we are referring to the Merge-Tree, the parent is
the next node toward the merge into the master branch, or the root of
the tree. When referring to the DAG, the parent relationship is in the
opposite direction, from the root toward the branch-point."

    The explanation of Algorithm 2 is, in its current state, generally
    very hard to follow. An example should be included, in a similar way
    as Figures 3, 4, and 5 helped to comprehend the first algorithm. The
    authors must seriously consider simplifying this explanation
    significantly, or the reader is simply lost. A step-by-step graphical
    explanation, using e.g. a flowchart or a diagram, would be much
    appreciated.

An example is included using the DAG in Figure 4. The steps are shown in
Table VIII. A high level description of the algorithm is provided in
paragraphs 3, with the subsequent paragraphs going into more detail.

    In Section C, the authors mention that “The left-most child of a merge
    is on the same branch as the merge.” But what does left-most mean in
    this context? What is left and right when considering the children of
    a node? This was not introduced.

This is no longer applicable. The 'left-most' was in reference to the
figure showing the visualization of the tree in the plugin. The section
has since been removed.

    Evaluation
    Figure 6 would benefit from a frame, or any other clearer visual cue
    that its text is not part of the main text, such as a different
    background color. Right now it merges (no pun intended) with the rest
    of the text.

A bounding box has been placed around the figure.

    In Section IV.A, the authors mention: “…compared the number of commits
    according to the log with the number of commits in the merge tree
    rooted in this commit.” However, it is not clear what “this commit”
    means, as there is no mention to a specific commit being analyzed.
    This is a little confusing and must be clarified. The way I
    understand, what the authors mean is that for each merge commit (mc),
    they (1) extract mc’s log message, (ii) extract from the log message
    the number of commits that were merged in mc, and (3) compare this
    number to the number of commits on mc’s merge-tree.

We replaced the section with a slight re-wording:

"
We extracted merges by Linus Torvalds using the command
git log --merges --author=’Torvalds’ and
compared the number of commits stated in the log message
with the number of commits in the Merge-Tree. We also
used the summaries of the commits found in the merge
(not necessarily all—see above) to make sure those commits
were in their corresponding Merge-Tree.
"

Basically, we look at the commit log. If it says there are 24 commits
and there are 24 commits in the merge tree, it is correct. If there are
more or fewer, then it is incorrect.

    In Section IV.B the authors suddenly mention a bitbucket limitation,
    but so far nothing has been introduced or explained about why the
    proposed technique is dependent on bitbucket. Also, the authors
    mention test cases, but don’t describe them concretey and
    unambiguously, using “…, and other various tests.” For reproducibility
    it would be preferred that the test suite is at least quickly
    described in full (if it is too long, then it should be provided as
    supplemental material to be available to the reader on demand). I
    recommend a rewrite of Section IV.B into a more concise, objective
    description of the tests and results.

Not applicable. This section has been removed, and the algorithm
generalization has been made into a discussion piece. The references to
bitbucket have been removed, as they are not necessary anymore.

    In Section V, in the sentence “These visualizations can be…”, what are
    these visualizations? No visualizations have been introduced so far.
    Also, the reference for the Reingold-Tilford tree should be shown in
    the first occurrence.

This part has been removed, instead opting to go directly into the
original use-cases that we were designing for.

"
The goal of Linvis is to simplify the navigation of the kernel
commit information, specifically focusing on merges. This is
done by leveraging the Merge-Tree to inspect how commits
are merged on the path to the master branch of the repository.

A. Use cases
"

    In Section 6, Reingold-Tilford tree, the authors describe that
    clicking a node will take me to that commit’s page, but I could not
    reproduce that behavior on the web application. Clicking a node simply
    centers that node on the screen.

Clicking a node will select that node (and center it), but not navigate to it.
Navigation requires clicking on the title that gets updated.

    Bubble Tree: there is now intermediate way of zooming in this
    representation, only from the root to the leaves and vice-versa. For
    big trees, this makes it very hard to navigate the tree. From an
    overview, the leaves are simply too small to be chosen (or clicked on)
    effectively and with accuracy. The navigation of the tree in its
    current state is not very useful.

Clicking on other circles will select that circle. Clicking on an
already selected circle will take you to the root. The interface was
updated slightly to increase the size of the border when the cursor is
hovering over a circle to indicate this.

For incredibly large trees, this is true, and I do not have ideas for
fixing this. Most trees are small though, containing less than 100
nodes.

    Although the two different tree visualizations are described, it is
    not clear how they can be actually used in order to extract new
    knowledge from the data. The authors are advised to add at least one
    interesting use case/example where the reader can follow,
    step-by-step, how the visualizations can be used to obtain insight
    from the data. Since the visualizations themselves are not novel, the
    novel use in this application context must be better motivated.

The new information is showing how a commit is merged into the master
branch of a repository, using different visualization metaphors. This
gives a person choices depending on which metaphor they like.

    Discussion
    In Section VIII, I agree that the proposed tool is more concise in
    showing the information than Gitk, however the practical use of the
    tool is still unclear. In order to replace the use that Gitk offers,
    the interface should be improved to include easier navigation and more
    information about each commit that is shown. In its current state, we
    see the nodes but it is very hard to have any kind of extra
    information about which are the commits we are seeing. The authors
    mention that “With Linvis we are able to immediately see what section
    of the kernel these commits and merges pertain to” but just by looking
    at Figure 18 that is not the case. In the paragraph starting on Page
    12, Line 19, the author talks about files that are nowhere to be
    found; maybe there is a missing figure to explain this point?

The discussion has shifted focus from the tool toward the results of the
study. For the specific use-cases described in section IV Linvis is
better than Gitk, but Linvis will never replace gitk.

    General Style Notes and Typos

Thank you, these have been addressed.

Responses to Reviewer 2:

    The focus of this paper is about two things notably the merge tree
    algorithm and visualizations. The merge tree algorithm part is
    impressive, while the visualizations are weak and not particularly
    novel. The paper shows a strong engineering effort. I am an advocate
    of this paper. Even though the algorithm is the stronger part, the
    contribution on visualization could be improved.

Thank you. The goal is to implement a visualization with the merge-tree
model so we can verify that it works, not to design a novel
visualization metaphor for trees.

    It is not clear how visualization will actually help in this project.
    There is a claim that it will, but no evidence supports that claim.
    Even with the visualizations that are presented it is not clear how to
    interpret them, interact with them, and what we can learn from them.
    An example that steps through the visualization techniques would be
    useful here, much like some of the panels described in the tool, to
    illustrate how effective the visualizations really are.

We conducted a user study to show that our tool/visualizations are able
to improve the correctness, accuracy, and timing when trying to
understand and summarize the events in the repository. See section VII.
We found that Linvis does impact comprehension and summarization.

    The LinVis tool is more like a search tool with visualizations as an
    added bonus. The visualizations almost feel like an afterthought. The
    paper does not describe in detail the justification of these
    visualization techniques over other ones. The visualizations
    themselves are not particularly novel. There is claims in the paper
    that the tool is effective, supports better navigation, and better
    explanation; there is simply no evidence. There is no evaluation of
    any kind. It is not clear how these claims are quantified. I
    understand an evaluation is a follow up project, but some details
    should be provided about how an evaluation is planned to be conducted.
    A few additional paragraphs to justify the visualization design and
    evaluation should suffice here.

From section V.C on page 7.

List tree:

"The nesting indicates the parent-child relationship. This visualization
is text-based enabling fast navigation to commits using the built-in
search in most web-browsers."

It allows users to quickly search for an exact commit, but isn't as good
as the others for visualizing the structure.

Reingold-Tilford Tree:

"is the classic tree visualization"

This is the tree diagram that computer scientists will draw if asked to
draw a tree structure. This tree is easy to understand and see how
commits are related and to which merge they belong, but is not great
with very wide trees.

Pack Tree:

"are useful for displaying an overview of large datasets. The initial
goal for the pack tree was for visualizing file systems, which are
similar to git repositories in that they are relatively shallow, but
very wide."

The pack tree visualization is designed to alleviate the remaining
issue.

    Missing related Work

Yes, added this to related work. There are likely many other
visualizations missing, but none that we found are designed with same
goals of showing how a commit is integrated into the master branch of a
repository.

    Regarding the term user

We are unaware of another concise term that is more gender-neutral than
the term "user" for describing people who use something. If there is a
term that you would prefer, we will be happy to looking into the
changes. For this reason, we retain the term "user" for general people
who will use the tool, regardless of intent.

We refer to the members of our user study as "participants".

    Minor Changes

Yes thank you. The structure of the paper has vastly changed since the
original paper was submitted. I believe that all of these have been
taken care of since then.